### 2019-10-23 Notes.Rmd
initial notes

---
title: "Notes"
author: "Yigit Ozan Berk"
date: "10/22/2019"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

Course Title

Data Science Specialization SwiftKey Capstone

Course Instructor(s)

Jeff Leek
Roger Peng
Brian Caffo
You are a data scientist now

The goal of this data science specialization has been to give you the basic skills involved with being a data scientist. The goal of this capstone is to mimic the experience of being a data scientist. As a practicing data scientist it is entirely common to get a messy data set, a vague question, and very little instruction on exactly how to analyze the data. Our goal is to give you that same experience but with added support in the form of forums, discussion with instructors, feedback from SwiftKey and Coursera engineers, and a structured problem to solve. We hope that you will take advantage of the opportunity this project affords for you to demonstrate your skills and creativity.

Course Tasks

This course will be separated into 8 different tasks that cover the range of activities encountered by a practicing data scientist. They mirror many of the skills you have developed in the data science specialization. The tasks are:

Understanding the problem
Data acquisition and cleaning
Exploratory analysis
Statistical modeling
Predictive modeling
Creative exploration
Creating a data product
Creating a short slide deck pitching your product
You will hear about each of these tasks over the course of the capstone.

Assessements and Grading

To successfully complete the capstone project, you must receive a passing grade on all of the following assignments:

Quiz 1: Getting Started
Milestone Report: exploratory analysis of the data set + evaluation of at least three classmate submissions
Quiz 2: Natural Language Processing I
Quiz 3: Natural Language Processing II
Final Project: your data product and a presentation describing your final data product + evaluation of at least three classmate submissions
The quizzes will be standard multiple choice quizzes. The other components are graded by peer evaluation.

Your final grade will be calculated as follows

Quiz 1 - 5%
Milestone Report - 20%
Quiz 2 - 10%
Quiz 3 -10%
Final Project -55%
Course dataset

This is the training data to get you started that will be the basis for most of the capstone. You must download the data from the link below and not from external websites to start.

https://d396qusza40orc.cloudfront.net/dsscapstone/dataset/Coursera-SwiftKey.zip
Later in the course you may use external data sets to augment your model as you see fit.

### 2019-10-25 ISLR_chp6.Rmd
ISLR chp6 lab subset selection

---
title: "ISLR Lab chp 6"
author: "Yigit Ozan Berk"
date: "10/25/2019"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# Chapter 6 Lab

## Best Subset Selection

```{r}
library(ISLR)
names(Hitters)
```

```{r}
dim(Hitters)
sum(is.na(Hitters$Salary))
```

```{r}
dat = na.omit(Hitters)
dim(dat)
```

```{r}
sum(is.na(dat))
```

The regsubsets() function (part of the `r leaps` library) performs best subset selection by identifying the best model that contains a given number of predictors, where best is quantified using RSS. The syntax is the same as for lm(). The summary() command outputs the best set of variables for each model size.
```{r}
install.packages("leaps")
library(leaps)
regfit.full = regsubsets(Salary~., dat)
summary(regfit.full)
```

an asterix indicates that a given variable is included in the corresponding model. For instance, this output indicates that the best two-variable model contans only Hits and CRBI. By default, regsubsets() only reports results up to the best eight-variable model. But the nvmax option can be used in order to return as many variables as are desired. Here we fit up to a 19-variable model.

```{r}
regfit.full = regsubsets(Salary ~., data = dat, nvmax = 19)
reg.summary = summary(regfit.full)
names(reg.summary)
```

the summary function also returns R^2 , RSS, adjusted R^2 C_p, and BIC. We can examine these to try to select the best overall model.

```{r}
reg.summary$rsq
```

we see that the R^2 statistic increases from 32%, when only one variable is included in the model, to almost 55%, when all variables are included. As expected, the R^2 statistic increases monotonically as more variables are included.

Plotting RSS, adjusted R^2, C_p and BIC for all of the models at once will help us decide which model to select. 
type = "l tells R to connect plotted points with lines

```{r}
par(mfrow = c(2,2))
plot(reg.summary$rss, xlab = "Number of variables", ylab = "RSS", type = "l")
plot(reg.summary$adjr2, xlab = "No of variables", ylab ="Adjusted RSq", type = "l")
which.max(reg.summary$adjr2)
#11
points(11, reg.summary$adjr2[11], col = "red", cex = 2, pch = 20)
```

in a similar fashion we can plot C_p and BIC statistics, and indicate the models with the smallest statistic using which.min()

```{r}
plot(reg.summary$cp, xlab= "no of variables", ylab = "Cp", type = "l")
which.min(reg.summary$cp)
#10
points(10, reg.summary$cp[10], col = "red", cwx = 2, pch = 20)
which.min(reg.summary$bic)
#6
plot(reg.summary$bic, xlab = "no of variables", ylab = "BIC", type = "l")
points(6, reg.summary$bic[6], col = "red", cex = 2, pch = 20)
```

the regsubsets() function has a built-in plot() command which can be used to display the selected variables for the best model with a given number of predictors, ranked according to the BIC, C_p, adjusted R^2, or AIC. To find out more about this function, type ?plot.regsubsets

```{r}

plot(regfit.full, scale = "r2")
plot(regfit.full, scale = "adjr2")
plot(regfit.full, scale = "Cp")
plot(regfit.full, scale = "bic")
```

the top row of each plot contains a black square for each variable selected according to the optimal model associated with that statistic. For instance, we see that several models share a BIC close to -150. However, the model with the lowest BIC is the six-variable model that contains only AtBat, Hits, Walks, CRBI, DivisionW, and PutOuts. We can use the coef() function to see the coefficient estimates associated with this model.
```{r}
coef(regfit.full, 6)
```

## Forward and Backward Stepwise Selection

We can also use the regsubsets() function to perform forward stepwise or backward stepwise selection, using the argument `r method = "forward"` or method = "backward"

```{r}
regfit.fwd = regsubsets(Salary~., data = dat, nvmax= 19, method = "forward")
summary(regfit.fwd)
regfit.bwd = regsubsets(Salary~., data = dat, nvmax = 19, method = "backward")
summary(regfit.bwd)
```

For instance, we see that using forward stepwise selection, the best one-variable mode contains only CRBI, and the best two-variable model additionally includes Hits. For this data, the best one-variable through six-variable models are each identical for best subset and forward selection. However, the best seven-variable models identified by forward stepwise selection, backward stepwise selection, and best subset selection are different.

```{r}
coef(regfit.full, 7)
coef(regfit.fwd, 7)
coef(regfit.bwd, 7)
```

## Choosing among models using Validation or CV

### 2019-10-29 ISLR_chp6.Rmd
lab chp 6

@@
 ## Choosing among models using Validation or CV
+
+```{r}
+library(ISLR)
+Hitters = na.omit(Hitters)
+set.seed(1)
+train = sample(c(TRUE, FALSE), nrow(Hitters), rep = TRUE)
+test = (!train)
+regfit.best = regsubsets(Salary~., data= Hitters[train, ], nvmax = 19)
+```
+
+we first make a model matrix from the test data
+
+```{r}
+test.mat = model.matrix(Salary~., data = Hitters[test,])
+```
+
+the model.matrix function is used in many regression packages for building an "X" matrix from data.
+
+now we run a loop, and for each size i, we extract the coefficients from regfit.best for the best model of that size, multiply them into the appropriate columns for the test model matrix to form the predictions, and compute the test MSE.
+
+```{r}
+val.errors = rep(NA, 19)
+for(i in 1:19){
+        coefi = coef(regfit.best, id = i)
+        pred = test.mat[ , names(coefi)]%*% coefi
+        val.errors[i] = mean((Hitters$Salary[test]- pred)^2)
+}
+```
+
+we find that the best model is the one that contains ten variables
+
+```{r}
+val.errors
+```
+
+```{r}
+which.min(val.errors)
+```
+
+```{r}
+coef(regfit.best, 10)
+```
+
+This was a little tedious, partly because there is no predict() method for regsubsets(). Since we will be using this function again, we can capture our steps above and write our own predict method.
+
+```{r}
+predict.regsubsets = function(object, newdata, id, ...) {
+        form = as.formula(object$call[[2]])
+        mat = model.matrix(form, newdata)
+        coefi = coef(object, id = id)
+        xvars = names(coefi)
+        mat[, xvars]%*% coefi
+}
+```
+
+our function pretty much mimics what we did above.
+the only complex part is how we extracted the formula used in the call to regsubsets().
+We demonstrate how we use this function below, when we do cross-validation.
+
+finally, we perform best subset selection on the full data set, and select the best ten-variable model.
+It is important that we make use of the full data set in order to obtain more accurate coefficient estimates.
+Note that we perform best subset selection on the full data set and select the best ten-variable model, rather than simply using the variables that were obtained from the training set, because the best ten-variable model on the full data set may differ from the corresponding model on the training set.
+
+```{r}
+regfit.best = regsubsets(Salary~., data = Hitters, nvmax = 19)
+coef(regfit.best, 10)
+```
+
+in fact, we see that the best ten-variable model on the full data set has a different set of variables than the best ten-variable model on the training set.
+
+we now try to choose among the models of different sizes using cross-validation.
+This approach is somewhat involved, as we must perform the best subset selection within each of the k training sets.
+Despite this, we see that with its clever subsetting syntax, R makes this job quite easy.
+First, we create a vector that alocates each observation to one of k = 10 folds, and we create a matrix in which we will store the results.
+
+```{r}
+k = 10
+set.seed(1)
+folds = sample(1:k, nrow(Hitters), replace = TRUE)
+# 1'den k'ya kadar olan rakamlarla toplam nrow(Hitters) uzunlugunda vektor yaratiyor. her rakam, k sayida cikacagi icin, CV'nin test kismi her sayi icin esit oluyor. EVREKA
+cv.errors = matrix(NA, k, 19, dimnames = list(NULL, paste(1:19)))
+```
+
+now we write a for loop that performs cross-validation.
+in the jth fold, the elements of folds that equal j are in the test set, and the remainder are in the training set.
+we make our predictions for each model size, (using our new predict() method), compute the test errors on the appropriate subset, and store them in the appropriate slot in the matrix cv.errors.
+
+```{r}
+for(j in 1:k) {
+        best.fit = regsubsets(Salary ~., data = Hitters[folds != j, ], nvmax = 19)
+        for(i in 1:19){
+                pred = predict(best.fit, Hitters[folds == j,], id = i)
+                cv.errors[j, i] = mean( (Hitters$Salary[folds == j]- pred)^2)
+        }
+}
+```
+
+this has given us a 10 x 19 matrix, of which the (i, j)th element corresponds to the test MSE for the ith cross-validation fold for the best j-variable model. 
+We use the apply() function to average over the columns of this matrix in order to obtain a vector for which the jth element is the cross-validation error for the j-variable model.
+
+```{r}
+mean.cv.errors = apply(cv.errors, 2, mean)
+mean.cv.errors
+```
+
+```{r}
+par(mfrow = c(1,1))
+plot(mean.cv.errors, type = "b")
+```
+
+we see that cross-validation selects an 11-variable model. we now perform best subset selection on the full data set in order to obtain the 11-variable model.
+
+```{r}
+reg.best = regsubsets(Salary~., data = Hitters, nvmax = 19)
+coef(reg.best, 11)
+```
+
+## Ridge Regression and the Lasso
+
+we will use the glmnet package
+
+```{r}
+install.packages("glmnet")
+library(glmnet)
+```
+
+This function has slightly different syntax from the other model-fitting functions that we have encountered thus far in the book.
+In particular, we must pass in an x matrix as well as a y vector, and we do not use the y~x syntax.
+We will now perform ridge regression and the lasso in order to predict Salary on the Hitters data.
+Before proceeding unsure that the missing values have been removed.
+
+```{r}
+x = model.matrix(Salary ~., Hitters)[,-1]
+y = Hitters$Salary
+```
+
+the model.matrix() function is particularly useful for creating x; not only does it produce a matrix corresponding to the 19 predictors but it also automatically transforms any qualitative variables into dummy variables.
+The latter property is important because glmnet() can only take numerical, quantitative inputs.
+
+### ridge regression
+
+glmnet() function has an `r alpha` argument that determines what type of model is fit. If alpha = 0, then a ridge regression model is fit, and if alpha = 1, then a lasso model is fit
+
+alpha = 0 -> ridge regression
+alpha = 1 -> lasso
+
+we first fit a ridge regression model
+
+```{r}
+grid = 10^seq(10, -2, length = 100)
+ridge.mod = glmnet(x, y, alpha = 0, lambda = grid)
+```
+
+by default, glmnet() function performs ridge regression for an automatically selected range of lambda values.
+Howeverm here we have chosen to implement the function over a grid of values ranging from 10^10 to 10^-2, essentially covering the full range of scenarios from the null model containing only the intercept, to the least squares fit.
+As we will see, we can also compute model fits for a particular value of lambda that is not one of the original grid values.
+Note that by default, the glmnet() function standardizes the variables so that they are on the same scale.
+You can turn off this setting using the argument standardize = FALSE.
+
+Associated with each value of labmda is a vector of ridge regression coefficients, stored in a matrix that can be accessed by coef().
+in this case, it is a 20x100 matrix with 20 rows(one for each predictor, plus an intercept) and 100 columns(one for each value of lambda).
+
+```{r}
+dim(coef(ridge.mod))
+```
+
+we expect the coefficient estimates to be much smaller, in terms of l_2 norm, when a large value of lambda is used, as compared to when a small value of lambda is used.
+these are the coefficients when lambda == 11,498, along with their l_2 norm:
+
+```{r}
+ridge.mod$lambda[50]
+```
+
+```{r}
+coef(ridge.mod)[,50]
+```
+
+```{r}
+sqrt(sum(coef(ridge.mod)[-1, 50]^2))
+#l_2 norm
+```
+
+in contrast, here are the coefficients when lambda == 705, along with their l_2 norm. note the much larger l_2 norm of the coefficients associated with this smaller value of lambda
+
+```{r}
+ridge.mod$lambda[60]
+```
+
+```{r}
+coef(ridge.mod)[,60]
+```
+
+```{r}
+sqrt(sum(coef(ridge.mod)[-1, 60]^2))
+```
+
+we can use predict function for a number of purposes.
+
+we can obtain the ridge regression coefficients for a new value of lambda, say 50
+
+```{r}
+predict(ridge.mod, s = 50, type = "coefficients")[1:20, ]
+```
+
+training / test samples for test error rate estimation
+
+```{r}
+set.seed(1)
+train = sample(1:nrow(x), nrow(x)/2)
+test = (-train)
+y.test = y[test]
+```
+
+fit the regression model on the training set, and evaluate its MSE on the test set, using lambda = 4.
+Note the use of predict function.
+This time we get predictions for a test set, by replacing typye = "coefficients" with the newx argument
+
+thresh = Convergence threshold for coordinate descent.
+Each inner coordinate-descent loop continues until the maximum change in the objective after any coefficient update is less than thresh times the null deviance.
+Defaults value is 1E-7.
+
+```{r}
+ridge.mod = glmnet(x[train,], y[train], alpha = 0, lambda = grid, thresh = 1e-12)
+#thresh nedir
+ridge.pred = predict(ridge.mod, s = 4, newx = x[test,])
+mean((ridge.pred - y.test)^2)
+```
+
+the test MSE is 101037.
+
+note that if we had instead simply fit a model with just an intercept, we would have predicted each test observation using the mean of the training observations.
+in that case, we could compute the test MSE like this:
+
+```{r}
+mean((mean(y[train]) - y.test)^2)
+```
+
+we could also get the same result by fitting a ridge regression model with a VERY large value of lambda. nore that 1e10 means 10^10
+
+```{r}
+ridge.pred = predict(ridge.mod, s = 1e10, newx = x[test, ])
+mean((ridge.pred - y.test)^2)
+```
+
+so fitting a ridge regression model with lambda ==4 leads to a much lower test MSE than fitting a model with just an intercept.
+we now check whether there is any benefit to performing ridge regression with lambda = 4 instead of just performing least squares regression.
+Recall that lest squares is simply ridge regression with lambda = 0.
+
+```{r}
+ridge.pred = predict(ridge.mod, s = 0, newx = x[test, ], exact = T, x = x[train, ], y = y[train])
+#here exact = T is for the glmnet to yield exact least squares coefficients when lambda is 0.
+mean((ridge.pred - y.test)^2)
+lm(y~x, subset = train)
+predict(ridge.mod, s= 0, exact = T, type = "coefficients", x = x[train, ], y = y[train])[1:20, ]
+```
+
+instead of arbitrarily choosing lambda = 4, we must use cross-validation to choose the tuning parameter lambda.
+we can do this using the built-in cross-validation function, cv.glmnet().
+by default, the function performs ten-fold CV, though this can be changed using the argument nfolds.
+
+```{r}
+set.seed(1)
+cv.out = cv.glmnet(x[train, ], y[train], alpha = 0)
+plot(cv.out)
+
+```
+
+```{r}
+bestlam = cv.out$lambda.min
+bestlam
+```
+
+Therefore, we see that the value of lambda that results in the smallest cross-validation error is 212. what is the test MSE associated with this value of lambda?
+
+```{r}
+ridge.pred = predict(ridge.mod, s = bestlam, newx = x[test, ])
+mean((ridge.pred - y.test)^2)
+```
+
+this represents a further improvement over the test MSE that we got using lambda = 4.
+finally, we refit our ridge regression model on the full data set, using the value of lambda chosen by CV, and examine the coefficient estimates.
+
+```{r}
+out = glmnet(x, y, alpha = 0)
+predict(out, type = "coefficients", s = bestlam)[1:20, ]
+```
+
+as expected, none of the coefficients are zero- ridge regression does not perform variable selection.
+
+### the lasso
+
+in order to fit a lasso model, we once again use glmnet() function; however, this time we use the argument alpha = 1. other than that change, we proceed just as we did in fitting a ridge model.
+
+```{r}
+lasso.mod = glmnet(x[train, ], y[train], alpha = 1, lambda = grid)
+plot(lasso.mod)
+```
+
+we can see from the plot that depending on the choice of tuning parameter, some of the coefficients will be exactly equal to zero. we now perform CV and compute the associated test error.
+
+```{r}
+set.seed(1)
+cv.out = cv.glmnet(x[train, ], y[train], alpha = 1)
+plot(cv.out)
+```
+
+```{r}
+bestlam = cv.out$lambda.min
+lasso.pred= predict(lasso.mod, s = bestlam, newx = x[test, ])
+mean((lasso.pred - y.test)^2)
+```
+
+substantially lower than test set MSE of the null model and of least squares, and very similar to the test MSE of ridge regression with lambda chosen by CV.
+
+however, lasso has substantial advantage over ridge regression: the resulting coefficient estimates are sparse.
+Here we see that 12 of the 19 coefficient estimates are exactly zero.
+So the lasso model with lambda chosen by CV contains only seven variables.
+
+```{r}
+out = glmnet(x, y, alpha =1 , lambda = grid)
+lasso.coef = predict(out, type = "coefficients", s = bestlam)[1:20, ]
+lasso.coef
+```
+
+## PCR and PLS Regression
+
+```{r}
+install.packages("pls")
+library(pls)
+```
+
+principal components regression (PCR) can be performed using the pcr() function.
+
+```{r}
+Hitters = na.omit(Hitters)
+set.seed(2)
+pcr.fit = pcr(Salary~., data = Hitters, scale = T, validation = "CV")
+```
+
+the syntax for the pcr function is similar to that for lm(), with a few additional options
+
+scale = TRUE for standardizing each predictor
+validation = "CV" causes pcr to compute ten-fold CV error for each possible value of M, the number of principal components used.
+the resulting fit can be examined using summary.
+
+```{r}
+summary(pcr.fit)
+```
+
+/the CV score is provided for each possible number of components, ranging from M = 0 onwards.
+
+Note that pcr() reports the ROOT MEAN SQUARED ERROR; in order to obtain the usual MSE, we must square this quantity.
+
+One can also plot the cross-validation scores using the validationplot() function. Using val.type = "MSEP" will cause the CV MSE to be plotted.
+
+```{r}
+validationplot(pcr.fit, val.type = "MSEP")
+```
+
+we see that hte smallest CV error occurs when M = 16 components are used.
+This is barely fewer than M = 19, which amounts to simply performing least squares, because when all of the components are used in PCR no dimension reduction occurs.
+However, from the plot we also see that the CV error is roughly the same when only one component is included in the model.
+This suggests that a model that uses just a small number of components might suffice.
+
+the summary() function also provides the percentage of variance explained in the predictors and in the reponse using different numbers of components.
+
+M = 1 only captures 38.31%
+M = 6 captures 88.63%
+
+we now perform PCR on training data, and evaluate its test set performance
+
+```{r}
+set.seed(1)
+train = sample(1:nrow(Hitters), nrow(Hitters)/2)
+test = (-train)
+x = model.matrix(Salary~., Hitters)[, -1]
+y = Hitters$Salary
+y.test = y[test]
+pcr.fit = pcr(Salary~., data = Hitters, subset = train, scale = T, validation = "CV")
+validationplot(pcr.fit, val.type = "MSEP")
+```
+
+now we find that the lowest CV error occurs when M = 7 components are used. We compute the test MSE as follows
+
+```{r}
+pcr.pred = predict(pcr.fit, x[test, ], ncomp = 7)
+mean((pcr.pred - y.test)^2)
+```
+
+this test MSE is competitive with the results obtained using ridge regression and the lasso.
+However, as a result of the way PCR is implemented, the final model is more difficult to interpret because it does not perform any kind of variable selection or even directly produce coefficient estimates.
+
+Finally, we fit PCR on the full data set, using M = 7, the number of components identified by CV.
+
+```{r}
+pcr.fit = pcr(y ~ x, scale = T, ncomp = 7)
+summary(pcr.fit)
+```
+
+### Partial Least Squares
+
+plsr() function in pls library
+
+```{r}
+set.seed(1)
+pls.fit = plsr(Salary ~., data = Hitters, subset = train, scale = T, validation = "CV")
+summary(pls.fit)
+```
+
+```{r}
+validationplot(pls.fit, val.type = "MSEP")
+```
+
+the lowest CV error occurs when only M = 2 pls directions are used. We now evaluate the corresponding test set MSE.
+
+```{r}
+pls.pred = predict(pls.fit, x[test, ], ncomp = 2)
+mean((pls.pred - y.test)^2)
+```
+
+the test MSE is comparable to, but slightly higher than, the test MSE obtained using ridge regression, the lasso, and PCR.
+
+Finally, we perform PLS using the full data set, using M= 2, the number of components identified by CV.
+
+```{r}
+pls.fit = plsr(Salary~., data = Hitters, scale = T, ncomp = 2)
+summary(pls.fit)
+```
+
+Notice the percentage of variance in Salary that the two-component PLS fit explains 46.40%, is almost as much as that explained using the final seven-component model PCR fit, 46.69%.
+This is because PCR only attempts to maximize the amount of variance explained in the predictors, while PLS searches for directions that explain variance in both the predictors and the response.
+
+# Exercises

### 2019-11-01 Notes.Rmd
week1 notes

@@
 Later in the course you may use external data sets to augment your model as you see fit.
+
+## Regular exp notes
+
+whitespace word boundaries
+
+sets of literals
+
+beginnig and end of a line
+
+alternatives (word1 or word2)
+
+- Start of a line
+"^i think"
+
+- end of the line
+"morning$"
+
+- sets of characters
+"[Bb][Uu][Ss][Hh]"
+all case alternatives for the word in e.g. BUSH, BuSH, Bush, bush, bUSH
+
+- all letters from a to z (uppercase or lower case)or numbers from 0 to 9, in the beginning of a line
+"^[0-9][a-zA-Z]"
+
+- when you use ^ at the beginning of a character class, it is a metacharacter indicating logical NOT operator.
+"[^?.]$"  means = any lines NOT ending with a question mark or a period.
+
+- "any" metacharacter
+"9.11" means= 9 followed by any character followed by 11
+
+- "or" metacharacter
+"flood|fire|coldfire" will match flood or fire or coldfire
+
+- you can combine
+"^[Gg]ood|[Bb]ad" will match Good/good in the beginning of a line or Bad/bad anywhere
+
+"^([Gg]ood|[Bb]ad)" will match Good/good or Bad/bad in the beginning of a line
+
+- "optional" metacharacter
+"[Gg]eorge( [Ww]\.)? [Bb]ush" will match where whitespace + [Ww] + . is optional
+
+- backslash escapes metacharacter so . means any character, \. means period
+
+- repetitions
+"*" means repeat any number of times, including none
+"+" means at least one
+
+"(.*)" will match any character of any number

### 2019-11-01 preproc.Rmd
week1 notes

---
title: "Main1"
author: "Yigit Ozan Berk"
date: "11/1/2019"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# Data import

```{r}
install.packages("tm")
install.packages("slam")
library(slam)
library(tm)

```

```{r}
con <- file("./en_US.twitter.txt", "r") 
twit = readLines(con, skipNul = T)
close(con)

```

```{r}
set.seed(1)
x = rbinom(2360148, size = 1, prob = .7)
#70% of data for training, 30% for testing
train = twit[x]
test = twit[!x]
rm(twit)
```

### 2019-11-01 Notes.Rmd
week 1 quiz done

@@
 "(.*)" will match any character of any number
+
+"[0-9]+ (.*)[0-9]+" will match, at least one number, followed by whitespace and any number of characters and at least one number again.
+
+- you can quantify the number of repetitions (min and max) you want with the curly brackets{}
+
+"[Bb]ush( +[^ ]+ +){1,5} debate" will match at least one space, followed by something that's not a space, followed by at least one space, and we want to see that between one and five times = we want to see something like space word space.
+
+{m,n} means at least m but not more than n matches
+{m} means exactly m matches
+{m,} means at least m matches
+
+- in most implementations of regular expressions, the parentheses not only limit the scope of alternatives divided by a "|", but also can be used to "remember" text matched by the subexpression enclosed
+we refer to the matched text with \1, \2, etc..
+
+" +([a-zA-Z]+) +\1 +" will match a space, followed by (at least one characters) followed by at least one space, followed by exactly the same match as before.
+ex: "............. so so ......."
+
+- * is greedy. it takes the longest possible string. you can use the ? to turn the greediness off.

### 2019-11-01 preproc.Rmd
week 1 quiz done

@@
 
+## twitter files
 ```{r}
@@
 close(con)
-
+y = NULL
+for(i in 1:length(twit)) {y[i] = nchar(twit[[i]])}
+summary(y)
 ```
@@
 ```
+
+## blogs
+
+```{r}
+con <- file("./en_US.blogs.txt", "r") 
+blogs = readLines(con, skipNul = T)
+close(con)
+
+y = NULL
+for(i in 1:length(blogs)) {y[i] = nchar(blogs[[i]])}
+summary(y)
+
+set.seed(123)
+x = rbinom(899288, size = 1, prob = .7)
+#70% of data for training, 30% for testing
+train = blog[x]
+test = blog[!x]
+```

### 2019-11-05 Notes.Rmd
quanteda startover

@@
 - * is greedy. it takes the longest possible string. you can use the ? to turn the greediness off.
+
+packages to use
+
+quanteda
+readtext
+spacyr
+
+```{r}
+install.packages("quanteda")
+install.packages("readtext")
+install.packages("spacyr")
+
+```
+
+# Loading in files using the readtext package
+```{r}
+require(readtext)
+
+# Twitter json
+dat_json <- readtext("~/Dropbox/QUANTESS/social media/zombies/tweets.json")
+corp_twitter <- corpus(dat_json)
+summary(corp_twitter, 5)
+# generic json - needs a textfield specifier
+dat_sotu <- readtext("~/Dropbox/QUANTESS/Manuscripts/collocations/Corpora/sotu/sotu.json",
+                  textfield = "text")
+summary(corpus(dat_sotu), 5)
+# text file
+dat_txtone <- readtext("~/Dropbox/QUANTESS/corpora/project_gutenberg/pg2701.txt", cache = FALSE)
+summary(corpus(dat_txtone), 5)
+# multiple text files
+dat_txtmultiple1 <- readtext("~/Dropbox/QUANTESS/corpora/inaugural/*.txt", cache = FALSE)
+summary(corpus(dat_txtmultiple1), 5)
+# multiple text files with docvars from filenames
+dat_txtmultiple2 <- readtext("~/Dropbox/QUANTESS/corpora/inaugural/*.txt",
+                             docvarsfrom = "filenames", sep = "-",
+                             docvarnames = c("Year", "President"))
+summary(corpus(dat_txtmultiple2), 5)
+# XML data
+dat_xml <- readtext("~/Dropbox/QUANTESS/quanteda_working_files/xmlData/plant_catalog.xml",
+                  textfield = "COMMON")
+summary(corpus(dat_xml), 5)
+# csv file
+write.csv(data.frame(inaug_speech = texts(data_corpus_inaugural),
+                     docvars(data_corpus_inaugural)),
+          file = "/tmp/inaug_texts.csv", row.names = FALSE)
+dat_csv <- readtext("/tmp/inaug_texts.csv", textfield = "inaug_speech")
+summary(corpus(dat_csv), 5)
+```

### 2019-11-05 preproc2.Rmd
quanteda startover

---
title: "PreProc2"
author: "Yigit Ozan Berk"
date: "11/5/2019"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# init

```{r, init}
library(quanteda)
library(readtext)
library(spacyr)
```

```{r, numwords}

##orig.wd <- getwd()
##setwd("../final/en_US")
numwords <- system("wc -w *.txt", intern=TRUE)  # intern=TRUE to return output  
numlines <- system("wc -l *.txt", intern=TRUE)
numbytes <- system("wc -c *.txt", intern=TRUE)
# longest <- system("wc -L *.txt", intern=TRUE) # not working
##setwd(orig.wd)  # return to original working dir, ie. the parent of /final

# number of words for each dataset
blog.numwords <- as.numeric(gsub('[^0-9]', '', numwords[1]))
news.numwords <- as.numeric(gsub('[^0-9]', '', numwords[2]))
twit.numwords <- as.numeric(gsub('[^0-9]', '', numwords[3]))

# number of lines for each dataset
blog.numlines <- as.numeric(gsub('[^0-9]', '', numlines[1]))
news.numlines <- as.numeric(gsub('[^0-9]', '', numlines[2]))
twit.numlines <- as.numeric(gsub('[^0-9]', '', numlines[3]))

# number of bytes for each dataset
blog.numbytes <- as.numeric(gsub('[^0-9]', '', numbytes[1]))
news.numbytes <- as.numeric(gsub('[^0-9]', '', numbytes[2]))
twit.numbytes <- as.numeric(gsub('[^0-9]', '', numbytes[3]))
```

# data import
```{r}
dat.twit <- readtext("./en_US.twitter.txt", cache = FALSE)
# summary(corpus(dat.twit), 5)
# or

con <- file("./en_US.twitter.txt", "r") 
twit = readLines(con, skipNul = T)
close(con)

#takin 15% of data for memory reasons
set.seed(12345)
x = sample(2360148, 350000, replace = F)

train = twit[x]

myN <- corpus(train)
rm(con, train, twit, x)
summary(myN, 5)
```

### 2019-11-06 preproc2.Rmd
week 2 bigrams and trigrams

@@
 
-```{r, init}
+```{r, init, eval=FALSE}
 library(quanteda)
@@
 library(spacyr)
+library(ggplot2)
+library(wordcloud)
 ```
@@
 # data import
-```{r}
+```{r, dataimport, cache = TRUE}
 dat.twit <- readtext("./en_US.twitter.txt", cache = FALSE)
@@
 ```
+
+```{r}
+texts(myN)[2]
+texts(myN)[4]
+texts(myN)[100]
+texts(myN)[1000]
+```
+
+## Data-Feature Matrix
+
+```{r}
+unigram = tokens(myN, remove_numbers = T, remove_punct = T)
+ngramTable
+```
+
+```{r}
+ngram1 <- tokens(myN[1],n=1)
+ngramTable <- as.data.frame(table(ngram1))
+# count the number of distinct 1-grams, matching Types for BNP text
+nrow(ngramTable)
+#1125
+# count total number of 1-grams (i.e. sum the frequencies), matching Tokens
+sum(ngramTable$Freq)
+#3280
+# print most frequent 1-grams
+tail(ngramTable[order(ngramTable$Freq),])
+```
+
+data feature matrix
+
+dat.dfm <- dfm(dat, ngrams=ng, toLower = FALSE, removePunct = FALSE, what = "fasterword", verbose = FALSE)
+```{r}
+dfm.twit = dfm(myN, remove_numbers = T, remove_punct = T)
+```
+
+top frequent words
+
+```{r}
+topfeatures(dfm.twit, 20)
+#dfm without stopwords
+dfm.twit = dfm(myN, remove_numbers = T, remove_punct = T,
+               remove = stopwords("english"))
+top20 = topfeatures(dfm.twit, 20)
+topfeatures(dfm.twit, 20)
+```
+
+```{r}
+top100 = topfeatures(dfm.twit, 100)
+topfeatures(dfm.twit, 100)
+```
+
+creating a wordcloud
+```{r}
+library(wordcloud)
+set.seed(100)
+textplot_wordcloud(dfm.twit, min_count = 3000, random_order = FALSE,
+                   rotation = .25,
+                   color = RColorBrewer::brewer.pal(8, "Dark2"))
+```
+
+```{r}
+qplot(y = top20, x = names(top20))
+```
+
+bigrams
+```{r}
+twit.bigram <- tokens(myN[1],n=2, remove_numbers = T, remove_punc = T)
+twit.bigram
+```
+
+```{r}
+dfm.twit.bigram <- dfm(myN, ngrams=2, remove_punct = T, remove_numbers = T, concatenator = " ", remove = stopwords("english"), verbose = T)
+dfm.twit.bigram[1:5, 1:5]
+```
+
+docfreq function
+
+(weighted) document frequency of a feature.
+
+```{r}
+freqs.twit = docfreq(dfm.twit.bigram)
+topfeatures(dfm.twit.bigram, 50)
+```
+
+trigrams
+
+```{r}
+dfm.twit.trigram <- dfm(myN, ngrams=3, remove_punct = T, remove_numbers = T)
+dfm.twit.trigram[1:5, 1:5]
+```
+
+```{r}
+topfeatures(dfm.twit.trigram, 50)
+```

### 2019-11-09 Notes.Rmd
week 2 bigrams and trigrams
milestone rep published

@@
 install.packages("spacyr")
-
+install.packages("wordcloud")
 ```
@@
 ```
+
+example
+
+```{r}
+ltcorpus = NULL
+ltcorpus[1] = "SOS buy the book EOS"   
+ltcorpus[2] = "SOS buy the book EOS"   
+ltcorpus[3] = "SOS buy the book EOS"   
+ltcorpus[4] = "SOS buy the book EOS"   
+ltcorpus[5] = "SOS sell the book EOS"
+ltcorpus[6] = "SOS buy the house EOS"
+ltcorpus[7] = "SOS buy the house EOS"
+ltcorpus[8] = "SOS paint the house EOS"
+```
+
+Step 1. i. Unigram, Bigram and Trigram counts
+```{r}
+## Returns a named vector of n-grams and their associated frequencies
+## extracted from the character vector dat.
+##
+## ng - Defines the type of n-gram to be extracted: unigram if ng=1,
+##      bigram if ng=2, trigram if n=3, etc.
+## dat - Character vector from which we want to get n-gram counts.
+## ignores - Character vector of words (features) to ignore from frequency table
+## sort.by.ngram - sorts the return vector by the names
+## sort.by.freq - sorts the return vector by frequency/count
+getFreqs = function(dat, ng) {
+        dat.dfm = dfm(dat, ngrams = ng, remove_punct = T, remove_numbers = T,
+                      remove = stopwords("english"))
+        dat.freq = docfreq(dat.dfm)
+        return(dat.freq)
+}
+
+getNgramFreqs <- function(ng, dat, ignores=NULL,
+                          sort.by.ngram=TRUE, sort.by.freq=FALSE) {
+    # http://stackoverflow.com/questions/36629329/
+    # how-do-i-keep-intra-word-periods-in-unigrams-r-quanteda
+    if(is.null(ignores)) {
+        dat.dfm <- dfm(dat, ngrams=ng, remove_punct = T, remove_numbers = T, 
+                       remove = stopwords("english"), 
+                       what = "fastestword")
+    } else {
+            dat.dfm <- dfm(dat, ngrams=ng, remove_punct = T, remove_numbers = T,
+                           remove = stopwords("english"), 
+                           ignoredFeatures = ignores, what = "fastestword")
+    }
+    rm(dat)
+    # quanteda docfreq will get the document frequency of terms in the dfm
+    ngram.freq <- docfreq(dat.dfm)
+    if(sort.by.freq) { ngram.freq <- sort(ngram.freq, decreasing=TRUE) }
+    if(sort.by.ngram) { ngram.freq <- ngram.freq[sort(names(ngram.freq))] }
+    rm(dat.dfm)
+
+    return(ngram.freq)
+}
+
+## Returns a 2 column data.table. The first column: ngram, contains all the
+## unigrams, bigrams, or trigrams in the corpus depending on whether
+## ng = 1, 2, or 3 respectively. The second column: freq, contains the
+## frequency or count of the ngram found in linesCorpus.
+##
+## ng - Defines the type of n-gram to be extracted: unigram if ng=1,
+##      bigram if ng=2, trigram if n=3, etc.
+## linesCorpus - character vector: each element is a line from a corpus file
+## prefixFilter - character vector: If not NULL, tells the function to return
+##                only rows where the ngram column starts with prefixFilter.
+##                If NULL, returns all the ngram and count rows.
+
+getTables = function(dat, ng) {
+        ngrams = getFreqs(dat = dat, ng = ng)
+        ngrams_dt = data.table(ngram = names(ngrams), freq = ngrams)
+        return(ngrams_dt)
+}
+
+getNgramTables <- function(ng, linesCorpus, prefixFilter=NULL) {
+    ngrams <- getNgramFreqs(ng, linesCorpus)
+    ngrams_dt <- data.table(ngram=names(ngrams), freq=ngrams)
+#    if(length(grep('^SOS', ngrams_dt$ngram)) > 0) {
+ #       ngrams_dt <- ngrams_dt[-grep('^SOS', ngrams_dt$ngram),]
+  #  }
+   # if(!is.null(prefixFilter)) {
+    #    regex <- sprintf('%s%s', '^', prefixFilter)
+    #    ngrams_dt <- ngrams_dt[grep(regex, ngrams_dt$ngram),]
+   # }
+   # 
+    return(ngrams_dt)
+}
+
+unigs <- getNgramTables(1, ltcorpus)
+bigrs <- getNgramTables(2, ltcorpus)
+trigs <- getNgramTables(3, ltcorpus)
+unigs; bigrs; trigs
+```
+
+Step 1. ii. Selecting bigram and trigram discounts
+
+For this example, we’ll use γ2=γ3=0.5
+ for the purpose of illustration.
+In practice, these values would be obtained by cross-validation.
+A great treatment of cross-validation can be found in Chapter 5 of this (free) book [10] which are discussed by the authors in these three videos: [11], [12], and [13].
+
+Step 2. Select Bigram Prefix of Word to be Predicted
+
+For this example, we’ll select the bigram: sell the
+
+Step 3. Calculate Probabilities of Words Completing Observed Trigrams
+
+```{r}
+gamma2 <- 0.5  # bigram discount
+gamma3 <- 0.5  # trigram discount
+bigPre <- 'sell_the'
+
+## Returns a two column data.frame of observed trigrams that start with the
+## bigram prefix (bigPre) in the first column named ngram and
+## frequencies/counts in the second column named freq. If no observed trigrams
+## that start with bigPre exist, an empty data.frame is returned.
+##
+## bigPre -  single-element char array of the form w2_w1 which are the first 
+##           two words of the trigram we are predicting the tail word of
+## trigrams - 2 column data.frame or data.table. The first column: ngram,
+##            contains all the trigrams in the corpus. The second column:
+##            freq, contains the frequency/count of each trigram.
+getObsTrigs <- function(bigPre, trigrams) {
+    trigs.winA <- data.frame(ngrams=vector(mode = 'character', length = 0),
+                             freq=vector(mode = 'integer', length = 0))
+    regex <- sprintf("%s%s%s", "^", bigPre, "_")
+    trigram_indices <- grep(regex, trigrams$ngram)
+    if(length(trigram_indices) > 0) {
+        trigs.winA <- trigrams[trigram_indices, ]
+    }
+
+    return(trigs.winA)
+}
+
+## Returns a two column data.frame of observed trigrams that start with bigram
+## prefix bigPre in the first column named ngram and the probabilities
+## q_bo(w_i | w_i-2, w_i-1) in the second column named prob calculated from
+## eqn 12. If no observed trigrams starting with bigPre exist, NULL is returned.
+##
+## obsTrigs - 2 column data.frame or data.table. The first column: ngram,
+##            contains all the observed trigrams that start with the bigram
+##            prefix bigPre which we are attempting to the predict the next
+##            word of in a give phrase. The second column: freq, contains the
+##            frequency/count of each trigram.
+## bigrs - 2 column data.frame or data.table. The first column: ngram,
+##         contains all the bigrams in the corpus. The second column:
+##         freq, contains the frequency/count of each bigram.
+## bigPre -  single-element char array of the form w2_w1 which are first two
+##           words of the trigram we are predicting the tail word of
+## triDisc - amount to discount observed trigrams
+getObsTriProbs <- function(obsTrigs, bigrs, bigPre, triDisc=0.5) {
+    if(nrow(obsTrigs) < 1) return(NULL)
+    obsCount <- filter(bigrs, ngram==bigPre)$freq[1]
+    obsTrigProbs <- mutate(obsTrigs, freq=((freq - triDisc) / obsCount))
+    colnames(obsTrigProbs) <- c("ngram", "prob")
+
+    return(obsTrigProbs)
+}
+
+obs_trigs <- getObsTrigs(bigPre, trigs)  # get trigrams and counts
+# convert counts to probabilities
+qbo_obs_trigrams <- getObsTriProbs(obs_trigs, bigrs, bigPre, gamma3)
+qbo_obs_trigrams
+```
+
+Step 4. Calculate Probabilities of Words Completing Unobserved Trigrams
+
+This is the the most complex step as it involves backing off to the bigram level. Here is a breakdown of the sub-steps for these calculations:
+
+Find all the words that complete unobserved trigrams. These are the words in the set w∈(wi−2,wi−1)
+ described earlier.
+Calculate α(wi−1)
+ from equation 14.
+Calculate qBO
+ for each bigram in the denominator of equation 17. using equation 10. if the bigram is observed or equation 16. if it is unobserved.
+Calculate α(wi−2,wi−1)
+ from equation 15.
+Calculate qBO(wi|wi−2,wi−1)
+ for each wi
+ from equation 17.
+Step 4. i. Find unobserved trigram tail words:
+
+```{r}
+## Returns a character vector which are the tail words of unobserved trigrams
+## that start with the first two words of obsTrigs (aka the bigram prefix).
+## These are the words w in the set B(w_i-2, w_i-1) as defined in the section
+## describing the details of equation 17.
+##
+## obsTrigs - character vector of observed trigrams delimited by _ of the form:
+##            w3_w2_w1 where w3_w2 is the bigram prefix
+## unigs - 2 column data.frame of all the unigrams in the corpus:
+##         ngram = unigram
+##         freq = frequency/count of each unigram
+getUnobsTrigTails <- function(obsTrigs, unigs) {
+    obs_trig_tails <- str_split_fixed(obsTrigs, "_", 3)[, 3]
+    unobs_trig_tails <- unigs[!(unigs$ngram %in% obs_trig_tails), ]$ngram
+    return(unobs_trig_tails)
+}
+
+unobs_trig_tails <- getUnobsTrigTails(obs_trigs$ngram, unigs)
+unobs_trig_tails
+```
+
+Step 4. ii. Calculate discounted probability mass at the bigram level α(wi−1)
+:
+
+The code below implements equation 14. to calculate α(wi−1)
+
+```{r}
+## Returns the total probability mass discounted from all observed bigrams
+## calculated from equation 14.  This is the amount of probability mass which
+## is redistributed to UNOBSERVED bigrams. If no bigrams starting with
+## unigram$ngram[1] exist, 0 is returned.
+##
+## unigram - single row, 2 column frequency table. The first column: ngram,
+##           contains the w_i-1 unigram (2nd word of the bigram prefix). The
+##           second column: freq, contains the frequency/count of this unigram.
+## bigrams - 2 column data.frame or data.table. The first column: ngram,
+##           contains all the bigrams in the corpus. The second column:
+##           freq, contains the frequency or count of each bigram.
+## bigDisc - amount to discount observed bigrams
+getAlphaBigram <- function(unigram, bigrams, bigDisc=0.5) {
+    # get all bigrams that start with unigram
+    regex <- sprintf("%s%s%s", "^", unigram$ngram[1], "_")
+    bigsThatStartWithUnig <- bigrams[grep(regex, bigrams$ngram),]
+    if(nrow(bigsThatStartWithUnig) < 1) return(0)
+    alphaBi <- 1 - (sum(bigsThatStartWithUnig$freq - bigDisc) / unigram$freq)
+
+    return(alphaBi)
+}
+
+unig <- str_split(bigPre, "_")[[1]][2]
+unig <- unigs[unigs$ngram == unig,]
+alpha_big <- getAlphaBigram(unig, bigrs, gamma2)
+alpha_big
+```
+
+Step 4. iii. Calculate backed off probabilities qBO
+ for bigrams
+
+The code below implements equation 10. to calculate qBO(wi|wi−1)
+ for observed bigrams and equation 16. for unobserved bigrams:
+
+```{r}
+## Returns a character vector of backed off bigrams of the form w2_w1. These 
+## are all the (w_i-1, w) bigrams where w_i-1 is the tail word of the bigram
+## prefix bigPre and w are the tail words of unobserved bigrams that start with
+## w_i-1.
+##
+## bigPre - single-element char array of the form w2_w1 which are first two
+##          words of the trigram we are predicting the tail word of
+## unobsTrigTails - character vector that are tail words of unobserved trigrams
+getBoBigrams <- function(bigPre, unobsTrigTails) {
+    w_i_minus1 <- str_split(bigPre, "_")[[1]][2]
+    boBigrams <- paste(w_i_minus1, unobsTrigTails, sep = "_")
+    return(boBigrams)
+}
+
+## Returns a two column data.frame of backed-off bigrams in the first column
+## named ngram and their frequency/counts in the second column named freq.
+## 
+## bigPre -  single-element char array of the form w2_w1 which are first two
+##           words of the trigram we are predicting the tail word of
+## unobsTrigTails - character vector that are tail words of unobserved trigrams
+## bigrs - 2 column data.frame or data.table. The first column: ngram,
+##         contains all the bigrams in the corpus. The second column:
+##         freq, contains the frequency/count of each bigram.
+getObsBoBigrams <- function(bigPre, unobsTrigTails, bigrs) {
+    boBigrams <- getBoBigrams(bigPre, unobsTrigTails)
+    obs_bo_bigrams <- bigrs[bigrs$ngram %in% boBigrams, ]
+    return(obs_bo_bigrams)
+}
+
+## Returns a character vector of backed-off bigrams which are unobserved.
+##
+## bigPre -  single-element char array of the form w2_w1 which are first two
+##           words of the trigram we are predicting the tail word of
+## unobsTrigTails - character vector that are tail words of unobserved trigrams
+## obsBoBigram - data.frame which contains the observed bigrams in a column
+##               named ngram
+getUnobsBoBigrams <- function(bigPre, unobsTrigTails, obsBoBigram) {
+    boBigrams <- getBoBigrams(bigPre, unobsTrigTails)
+    unobs_bigs <- boBigrams[!(boBigrams %in% obsBoBigram$ngram)]
+    return(unobs_bigs)
+}
+
+## Returns a dataframe of 2 columns: ngram and probs.  Values in the ngram
+## column are bigrams of the form: w2_w1 which are observed as the last
+## two words in unobserved trigrams.  The values in the prob column are
+## q_bo(w1 | w2) calculated from from equation 10.
+##
+## obsBoBigrams - a dataframe with 2 columns: ngram and freq. The ngram column
+##                contains bigrams of the form w1_w2 which are observed bigrams
+##                that are the last 2 words of unobserved trigrams (i.e. "backed
+##                off" bigrams). The freq column contains integers that are
+##                the counts of these observed bigrams in the corpus.
+## unigs - 2 column data.frame of all the unigrams in the corpus:
+##         ngram = unigram
+##         freq = frequency/count of each unigram
+## bigDisc - amount to discount observed bigrams
+getObsBigProbs <- function(obsBoBigrams, unigs, bigDisc=0.5) {
+    first_words <- str_split_fixed(obsBoBigrams$ngram, "_", 2)[, 1]
+    first_word_freqs <- unigs[unigs$ngram %in% first_words, ]
+    obsBigProbs <- (obsBoBigrams$freq - bigDisc) / first_word_freqs$freq
+    obsBigProbs <- data.frame(ngram=obsBoBigrams$ngram, prob=obsBigProbs)
+
+    return(obsBigProbs)
+}
+
+## Returns a dataframe of 2 columns: ngram and prob.  Values in the ngram
+## column are unobserved bigrams of the form: w2_w1.  The values in the prob
+## column are the backed off probability estimates q_bo(w1 | w2) calculated
+## from from equation 16.
+##
+## unobsBoBigrams - character vector of unobserved backed off bigrams
+## unigs - 2 column data.frame of all the unigrams in the corpus:
+##         ngram = unigram
+##         freq = frequency/count of each unigram
+## alphaBig - total discounted probability mass at the bigram level
+getQboUnobsBigrams <- function(unobsBoBigrams, unigs, alphaBig) {
+    # get the unobserved bigram tails
+    qboUnobsBigs <- str_split_fixed(unobsBoBigrams, "_", 2)[, 2]
+    w_in_Aw_iminus1 <- unigs[!(unigs$ngram %in% qboUnobsBigs), ]
+    # convert to data.frame with counts
+    qboUnobsBigs <- unigs[unigs$ngram %in% qboUnobsBigs, ]
+    denom <- sum(qboUnobsBigs$freq)
+    # converts counts to probabilities
+    qboUnobsBigs <- data.frame(ngram=unobsBoBigrams,
+                               prob=(alphaBig * qboUnobsBigs$freq / denom))
+
+    return(qboUnobsBigs)
+}
+
+bo_bigrams <- getBoBigrams(bigPre, unobs_trig_tails)  # get backed off bigrams
+# separate bigrams which use eqn 10 and those that use 16
+obs_bo_bigrams <- getObsBoBigrams(bigPre, unobs_trig_tails, bigrs)
+unobs_bo_bigrams <- getUnobsBoBigrams(bigPre, unobs_trig_tails, obs_bo_bigrams)
+# unobs_bo_bigrams = c("the_buy", "the_EOS", "the_paint", "the_sell", "the_the")
+# calc obs'd bigram prob's from eqn 10
+qbo_obs_bigrams <- getObsBigProbs(obs_bo_bigrams, unigs, gamma2) #ngram     probs
+# calc alpha_big & unobs'd bigram prob's from eqn 16             #the_house 0.3125
+unig <- str_split(bigPre, "_")[[1]][2]
+unig <- unigs[unigs$ngram == unig,]
+# distrib discounted bigram prob mass to unobs bigrams in prop to unigram ML
+qbo_unobs_bigrams <- getQboUnobsBigrams(unobs_bo_bigrams, unigs, alpha_big)
+qbo_bigrams <- rbind(qbo_obs_bigrams, qbo_unobs_bigrams)
+qbo_bigrams
+```
+
+Checking the Bigram Calculations
+
+Before doing the final calculations for the unobserved trigrams, let’s do a simple check on our calculations at the bigram level.
+In the previous table, all the bigrams except the_house are unobserved which means that if we sum all the unobserved bigram probabilities, we should get the total bigram discount which is α(wi−1)
+. As we see below, this looks like it checks out.
+
+```{r}
+unobs <- qbo_bigrams[-1,]
+sum(unobs$prob)
+```
+
+Step 4. iv. Calculate discounted probability mass at the trigram level α(wi−2,wi−1)
+
+The getAlphaTrigram function shown below implements equation 15. to compute the trigram discount. Here we use it to compute the trigram discount for qBO(house|sell,the)
+:
+
+```{r}
+## Returns the total probability mass discounted from all observed trigrams.
+## calculated from equation 14. This is the amount of probability mass which is
+## redistributed to UNOBSERVED trigrams. If no trigrams starting with
+## bigram$ngram[1] exist, 1 is returned.
+##
+## obsTrigs - 2 column data.frame or data.table. The first column: ngram,
+##            contains all the observed trigrams that start with the bigram
+##            prefix we are attempting to the predict the next word of. The 
+##            second column: freq, contains the frequency/count of each trigram.
+## bigram - single row frequency table where the first col: ngram, is the bigram
+##          which are the first two words of unobserved trigrams we want to
+##          estimate probabilities of (same as bigPre in other functions listed
+##          prior) delimited with '_'. The second column: freq, is the
+##          frequency/count of the bigram listed in the ngram column.
+## triDisc - amount to discount observed trigrams
+getAlphaTrigram <- function(obsTrigs, bigram, triDisc=0.5) {
+    if(nrow(obsTrigs) < 1) return(1)
+    alphaTri <- 1 - sum((obsTrigs$freq - triDisc) / bigram$freq[1])
+
+    return(alphaTri)
+}
+
+bigram <- bigrs[bigrs$ngram %in% bigPre, ]
+alpha_trig <- getAlphaTrigram(obs_trigs, bigram, gamma3)
+alpha_trig
+```
+
+Step 4. v. Calculate unobserved trigram probabilities qBO(wi|wi−2,wi−1)
+:
+
+```{r}
+## Returns a dataframe of 2 columns: ngram and prob.  Values in the ngram
+## column are unobserved trigrams of the form: w3_w2_w1.  The values in the prob
+## column are q_bo(w1 | w3, w2) calculated from equation 17.
+##
+## bigPre -  single-element char array of the form w2_w1 which are first two
+##           words of the trigram we are predicting the tail word of
+## qboObsBigrams - 2 column data.frame with the following columns -
+##                 ngram: observed bigrams of the form w2_w1
+##                 probs: the probability estimate for observed bigrams:
+##                        qbo(w1 | w2) calc'd from equation 10.
+## qboUnobsBigrams - 2 column data.frame with the following columns -
+##                   ngram: unobserved bigrams of the form w2_w1
+##                   probs: the probability estimate for unobserved bigrams
+##                          qbo(w1 | w2) calc'd from equation 16.
+## alphaTrig - total discounted probability mass at the trigram level
+getUnobsTriProbs <- function(bigPre, qboObsBigrams,
+                             qboUnobsBigrams, alphaTrig) {
+    qboBigrams <- rbind(qboObsBigrams, qboUnobsBigrams)
+    qboBigrams <- qboBigrams[order(-qboBigrams$prob), ]
+    sumQboBigs <- sum(qboBigrams$prob)
+    first_bigPre_word <- str_split(bigPre, "_")[[1]][1]
+    unobsTrigNgrams <- paste(first_bigPre_word, qboBigrams$ngram, sep="_")
+    unobsTrigProbs <- alphaTrig * qboBigrams$prob / sumQboBigs
+    unobsTrigDf <- data.frame(ngram=unobsTrigNgrams, prob=unobsTrigProbs)
+
+    return(unobsTrigDf)
+}
+
+qbo_unobs_trigrams <- getUnobsTriProbs(bigPre, qbo_obs_bigrams,
+                                       qbo_unobs_bigrams, alpha_trig)
+qbo_unobs_trigrams
+```
+
+Step 5. Select wi
+ with the highest qBO(wi|wi−2,wi−1)
+
+We’ve done all the calculations required to make our prediction. These are summarized in the table below:
+
+```{r}
+getPredictionMsg <- function(qbo_trigs) {
+    # pull off tail word of highest prob trigram
+    prediction <- str_split(qbo_trigs$ngram[1], "_")[[1]][3]
+    result <- sprintf("%s%s%s%.4f", "highest prob prediction is >>> ", prediction,
+                      " <<< which has probability = ", qbo_trigs$prob[1])
+    return(result)
+}
+
+qbo_trigrams <- rbind(qbo_obs_trigrams, qbo_unobs_trigrams)
+qbo_trigrams <- qbo_trigrams[order(-qbo_trigrams$prob), ]  # sort by desc prob
+out_msg <- getPredictionMsg(qbo_trigrams)
+out_msg
+```

### 2019-11-09 draft1.Rmd
milestone rep published

---
title: "SwiftKey Capstone Proj. Milestone Report"
author: "Yigit Ozan Berk"
date: "11/9/2019"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# Initiation
```{r}
library(quanteda)
library(readtext)
library(spacyr)
library(ggplot2)
library(wordcloud)
library(data.table)
```

# Downloading Data

The dataset can be downloaded from a link given in the course website. [https://d396qusza40orc.cloudfront.net/dsscapstone/dataset/Coursera-SwiftKey.zip]

The unzipped file contains a directory called final, then a subdirectory called en_US, which contains the texts.

There are 3 text files.
* en_US.blogs.txt - text from blog posts
* en_US.news.txt - text from news articles
* en_US.twitter.txt - tweets on Twitter

# Introduction

The goal here to display the initial data exploration, and show I am on the track to extablishing my algorithm.

```{r}
numwords <- system("wc -w *.txt", intern=TRUE)  # intern=TRUE to return output  
numlines <- system("wc -l *.txt", intern=TRUE)
numbytes <- system("wc -c *.txt", intern=TRUE)

# number of words for each dataset
blog.numwords <- as.numeric(gsub('[^0-9]', '', numwords[1]))
news.numwords <- as.numeric(gsub('[^0-9]', '', numwords[2]))
twit.numwords <- as.numeric(gsub('[^0-9]', '', numwords[3]))

# number of lines for each dataset
blog.numlines <- as.numeric(gsub('[^0-9]', '', numlines[1]))
news.numlines <- as.numeric(gsub('[^0-9]', '', numlines[2]))
twit.numlines <- as.numeric(gsub('[^0-9]', '', numlines[3]))

# number of bytes for each dataset
blog.numbytes <- as.numeric(gsub('[^0-9]', '', numbytes[1]))
news.numbytes <- as.numeric(gsub('[^0-9]', '', numbytes[2]))
twit.numbytes <- as.numeric(gsub('[^0-9]', '', numbytes[3]))

words = rbind(blog.numwords, news.numwords, twit.numwords)
lines = rbind(blog.numlines, news.numlines, twit.numlines)
bytes = rbind(blog.numbytes, news.numbytes, twit.numbytes)

data.frame(words = words, lines= lines, Mb = bytes/1000000, 
           row.names = c("blog", "news", "twit"))
```

For memory storage limitations, only %15 random entries from each dataset is used for calculations.

Twitter text
```{r}
con <- file("./en_US.twitter.txt", "r") 
twit = readLines(con, skipNul = T)
close(con)

#takin 15% of data for memory reasons
set.seed(1)
x = sample(2360148, 350000, replace = F)

train = twit[x]

TW <- corpus(train)
rm(con, train, twit, x)
summary(TW, 5)
```

Blog text
```{r}
con <- file("./en_US.blogs.txt", "r") 
blog = readLines(con, skipNul = T)
close(con)

#takin 15% of data for memory reasons
set.seed(12345)
x = sample(899288, 135000, replace = F)

train = blog[x]

BG <- corpus(train)
rm(con, train, blog, x)
summary(BG, 5)
```

News text

```{r}
# 1010242
con <- file("./en_US.news.txt", "r") 
news = readLines(con, skipNul = T)
close(con)

#takin 15% of data for memory reasons
set.seed(10000)
x = sample(1010242, 150000, replace = F)

train = news[x]

NS <- corpus(train)
rm(con, train, news, x)
summary(NS, 5)
```

# Visualize Word Frequency

```{r}
BG.uni = getTables(dat = BG, ng = 1)
set.seed(100)
textplot_wordcloud(BG.uni, min_count = 3000, random_order = FALSE,
                   rotation = .25,
                   color = RColorBrewer::brewer.pal(8, "Dark2"))
```

1. create frequency tables and n-grams
```{r}
getFreqs = function(dat, ng) {
        dat.dfm = dfm(dat, ngrams = ng, remove_punct = T, remove_numbers = T,
                      remove = stopwords("english"))
        dat.freq = docfreq(dat.dfm)
        dat.freq = dat.freq[sort(names(dat.freq))] 
        return(dat.freq)
}

getTables = function(dat, ng) {
        ngrams = getFreqs(dat = dat, ng = ng)
        ngrams_dt = data.table(ngram = names(ngrams), freq = ngrams)
        return(ngrams_dt)
}
```

```{r}
gamma2 <- 0.5  # bigram discount
gamma3 <- 0.5  # trigram discount

bigPre <- 'sell_the'

getObsTrigs <- function(bigPre, trigrams) {
    trigs.winA <- data.frame(ngrams=vector(mode = 'character', length = 0),
                             freq=vector(mode = 'integer', length = 0))
    regex <- sprintf("%s%s%s", "^", bigPre, "_")
    trigram_indices <- grep(regex, trigrams$ngram)
    if(length(trigram_indices) > 0) {
        trigs.winA <- trigrams[trigram_indices, ]
    }

    return(trigs.winA)
}

getObsTriProbs <- function(obsTrigs, bigrs, bigPre, triDisc=0.5) {
    if(nrow(obsTrigs) < 1) return(NULL)
    obsCount <- filter(bigrs, ngram==bigPre)$freq[1]
    obsTrigProbs <- mutate(obsTrigs, freq=((freq - triDisc) / obsCount))
    colnames(obsTrigProbs) <- c("ngram", "prob")

    return(obsTrigProbs)
}

obs_trigs <- getObsTrigs(bigPre, trigs)  # get trigrams and counts
# convert counts to probabilities
qbo_obs_trigrams <- getObsTriProbs(obs_trigs, bigrs, bigPre, gamma3)
qbo_obs_trigrams
```

```{r}
getUnobsTrigTails <- function(obsTrigs, unigs) {
    obs_trig_tails <- str_split_fixed(obsTrigs, "_", 3)[, 3]
    unobs_trig_tails <- unigs[!(unigs$ngram %in% obs_trig_tails), ]$ngram
    return(unobs_trig_tails)
}

unobs_trig_tails <- getUnobsTrigTails(obs_trigs$ngram, unigs)
unobs_trig_tails
```

### 2019-11-09 milestonerep.Rmd
milestone rep published

---
title: "SwiftKey Capstone Proj. Milestone Report"
author: "Yigit Ozan Berk"
date: "11/9/2019"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# Executive Summary

This is the first report for a data science project that involves word prediction using NLP. The ultimate goal for the project is to write an efficient algorithm that uses n-grams(script summaries used to tokenize the given script) to predict the next word that will appear after a given phrase.

In this report, I explore 3 text files - one from blogs, news, and twitter.

After background research, I have decided to use quanteda and data.table for my main calculations for better performance.

For initial visualisation, wordcloud package seemed most impressive.

# Initiation
```{r init}
library(quanteda)
library(spacyr)
library(ggplot2)
library(wordcloud)
library(data.table)
library(plotly)

# Below functions are used to tokenize the given script by the 'n-gram' value, 
# and return the frequency of each ngram
# what = "fasterword" or what = "fastestword" is not used due to incomplete cleaning measures
getFreqs = function(dat, ng) {
        dat.dfm = dfm(dat, ngrams = ng, remove_punct = T, remove_numbers = T,
                      remove = stopwords("english"))
        dat.freq = docfreq(dat.dfm)
        dat.freq = dat.freq[sort(names(dat.freq))] 
        return(dat.freq)
}

getTables = function(dat, ng) {
        ngrams = getFreqs(dat = dat, ng = ng)
        ngrams_dt = data.table(ngram = names(ngrams), freq = ngrams)
        return(ngrams_dt)
}
```

# Downloading Data

The dataset can be downloaded from a link given in the course website. [https://d396qusza40orc.cloudfront.net/dsscapstone/dataset/Coursera-SwiftKey.zip]

The unzipped file contains a directory called final, then a subdirectory called en_US, which contains the texts.

There are 3 text files.
* en_US.blogs.txt - text from blog posts
* en_US.news.txt - text from news articles
* en_US.twitter.txt - tweets on Twitter

# Introduction

The goal here to display the initial data exploration, and show I am on the track to extablishing my algorithm.

```{r exp}
numwords <- system("wc -w *.txt", intern=TRUE)  # intern=TRUE to return output  
numlines <- system("wc -l *.txt", intern=TRUE)
numbytes <- system("wc -c *.txt", intern=TRUE)

# number of words for each dataset
blog.numwords <- as.numeric(gsub('[^0-9]', '', numwords[1]))
news.numwords <- as.numeric(gsub('[^0-9]', '', numwords[2]))
twit.numwords <- as.numeric(gsub('[^0-9]', '', numwords[3]))

# number of lines for each dataset
blog.numlines <- as.numeric(gsub('[^0-9]', '', numlines[1]))
news.numlines <- as.numeric(gsub('[^0-9]', '', numlines[2]))
twit.numlines <- as.numeric(gsub('[^0-9]', '', numlines[3]))

# number of bytes for each dataset
blog.numbytes <- as.numeric(gsub('[^0-9]', '', numbytes[1]))
news.numbytes <- as.numeric(gsub('[^0-9]', '', numbytes[2]))
twit.numbytes <- as.numeric(gsub('[^0-9]', '', numbytes[3]))

words = rbind(blog.numwords, news.numwords, twit.numwords)
lines = rbind(blog.numlines, news.numlines, twit.numlines)
bytes = rbind(blog.numbytes, news.numbytes, twit.numbytes)

#summary
data.frame(words = words, lines= lines, Mb = bytes/1000000, 
           row.names = c("blog", "news", "twit"))
```

For memory storage limitations, only %15 random entries from each dataset is used for calculations.

Twitter text
```{r}
con <- file("./en_US.twitter.txt", "r") 
twit = readLines(con, skipNul = T)
close(con)

#takin 15% of data for memory reasons
set.seed(1)
x = sample(2360148, 350000, replace = F)

train = twit[x]

TW <- corpus(train)
rm(con, train, twit, x)
summary(TW, 5)
```

Blog text
```{r}
con <- file("./en_US.blogs.txt", "r") 
blog = readLines(con, skipNul = T)
close(con)

#takin 15% of data for memory reasons
set.seed(12345)
x = sample(899288, 135000, replace = F)

train = blog[x]

BG <- corpus(train)
rm(con, train, blog, x)
summary(BG, 5)
```

News text

```{r}

con <- file("./en_US.news.txt", "r") 
news = readLines(con, skipNul = T)
close(con)

#takin 15% of data for memory reasons
set.seed(10000)
x = sample(1010242, 150000, replace = F)

train = news[x]

NS <- corpus(train)
rm(con, train, news, x)
summary(NS, 5)
```

# Visualize Word Frequency

Here for each corpus file(BG, NS, TW), we create a data-feature matrix with quanteda::dfm() function. The selected features for the function are:
- remove_punct ; for cleaning punctuations from text
- remove_numbers ; for cleaning numbers from text
- remove = stopwords("english") ; for cleaning stopwords defined for english language by quanteda

After creating the data-feature matrix, a wordcloud visualisation is created with the wordcloud::textplot_wordcloud() function - for the words that appear more than 3000 times in total.

After wordcloud visualisation,

1.Blog text
```{r}
BG.uni = dfm(BG, ngrams = 1, remove_punct = T, remove_numbers = T,
                      remove = stopwords("english"))
set.seed(100)
textplot_wordcloud(BG.uni, min_count = 3000, random_order = FALSE,
                   rotation = .25,
                   color = RColorBrewer::brewer.pal(8, "Dark2"))
rm(BG.uni)
```

The top frequencies of the unique words as below:
```{r}
headBG = getTables(dat = BG, ng = 1)[order(freq, decreasing = T)][1:10,]
plot_ly(x = headBG$ngram, y = headBG$freq)
rm(headBG)
```

2.News text
```{r}
NS.uni = dfm(NS, ngrams = 1, remove_punct = T, remove_numbers = T,
                      remove = stopwords("english"))
set.seed(1)
textplot_wordcloud(NS.uni, min_count = 3000, random_order = FALSE,
                   rotation = .25,
                   color = RColorBrewer::brewer.pal(8, "Dark2"))
rm(NS.uni)
```

The top frequencies of the unique words as below:
```{r}
headNS = getTables(dat = NS, ng = 1)[order(freq, decreasing = T)][1:10,]
plot_ly(x = headNS$ngram, y = headNS$freq)
rm(headNS)
```

3. Twitter text
```{r}
TW.uni = dfm(TW, ngrams = 1, remove_punct = T, remove_numbers = T,
                      remove = stopwords("english"))
set.seed(12345)
textplot_wordcloud(TW.uni, min_count = 3000, random_order = FALSE,
                   rotation = .25,
                   color = RColorBrewer::brewer.pal(8, "Dark2"))
rm(TW.uni)
```

The top frequencies of the unique words as below:
```{r}
headTW = getTables(dat = TW, ng = 1)[order(freq, decreasing = T)][1:10,]
plot_ly(x = headTW$ngram, y = headTW$freq)
rm(headTW)
```

# Plans for the main algorithm

To make the algorithm into an app, first I need to establish the proper functions to create observed/unobserved trigram/bigram probability tables for a given phrase - using Katz Back-Off algorithm. After creating these functions, I can calculate the probabilities and find the highest-prob next word prediction for a given phrase.

For performance related issues, I will only use trigrams, bigrams, and unigrams.

Given a phrase, the algorithm will first check the trigram to find out most likely word. If there is no probable answer, it will check the bigram. If still no probable answer, it will finally check the unigram to predict the next word based on the most common single word in the corpus.

### 2019-11-09 preproc2.Rmd
milestone rep published

@@
 library(wordcloud)
+library(data.table)
 ```
@@
 ```{r}
-dfm.twit.bigram <- dfm(myN, ngrams=2, remove_punct = T, remove_numbers = T, concatenator = " ", remove = stopwords("english"), verbose = T)
+dfm.twit.bigram <- dfm(myN, ngrams=2, remove_punct = T, remove_numbers = T, remove = stopwords("english"), verbose = T)
 dfm.twit.bigram[1:5, 1:5]

### 2019-11-12 draft1.Rmd
full prediction algorithm

@@
 library(quanteda)
+library(dplyr)
+library(stringr)
 library(readtext)
@@
 }
+
 ```
 
+unigram, bigram, and trigrams
+
+```{r}
+unigs = getTables(BG, 1)
+bigrs = getTables(BG, 2)
+trigs = getTables(BG, 3)
+```
+
+Probs of words completing OBSERVED trigrams
 ```{r}
@@
 gamma3 <- 0.5  # trigram discount
+bigPre <- 'is_book'
 
-bigPre <- 'sell_the'
-
+## Returns a two column data.frame of observed trigrams that start with the
+## bigram prefix (bigPre) in the first column named ngram and
+## frequencies/counts in the second column named freq. If no observed trigrams
+## that start with bigPre exist, an empty data.frame is returned.
+##
+## bigPre -  single-element char array of the form w2_w1 which are the first 
+##           two words of the trigram we are predicting the tail word of
+## trigrams - 2 column data.frame or data.table. The first column: ngram,
+##            contains all the trigrams in the corpus. The second column:
+##            freq, contains the frequency/count of each trigram.
 getObsTrigs <- function(bigPre, trigrams) {
@@
 
+## Returns a two column data.frame of observed trigrams that start with bigram
+## prefix bigPre in the first column named ngram and the probabilities
+## q_bo(w_i | w_i-2, w_i-1) in the second column named prob calculated from
+## eqn 12. If no observed trigrams starting with bigPre exist, NULL is returned.
+##
+## obsTrigs - 2 column data.frame or data.table. The first column: ngram,
+##            contains all the observed trigrams that start with the bigram
+##            prefix bigPre which we are attempting to the predict the next
+##            word of in a give phrase. The second column: freq, contains the
+##            frequency/count of each trigram.
+## bigrs - 2 column data.frame or data.table. The first column: ngram,
+##         contains all the bigrams in the corpus. The second column:
+##         freq, contains the frequency/count of each bigram.
+## bigPre -  single-element char array of the form w2_w1 which are first two
+##           words of the trigram we are predicting the tail word of
+## triDisc - amount to discount observed trigrams
 getObsTriProbs <- function(obsTrigs, bigrs, bigPre, triDisc=0.5) {
@@
 
+Find UNOBSERVED trigram tail words
 ```{r}
+## Returns a character vector which are the tail words of unobserved trigrams
+## that start with the first two words of obsTrigs (aka the bigram prefix).
+## These are the words w in the set B(w_i-2, w_i-1) as defined in the section
+## describing the details of equation 17.
+##
+## obsTrigs - character vector of observed trigrams delimited by _ of the form:
+##            w3_w2_w1 where w3_w2 is the bigram prefix
+## unigs - 2 column data.frame of all the unigrams in the corpus:
+##         ngram = unigram
+##         freq = frequency/count of each unigram
 getUnobsTrigTails <- function(obsTrigs, unigs) {
@@
 unobs_trig_tails <- getUnobsTrigTails(obs_trigs$ngram, unigs)
-unobs_trig_tails
+head(unobs_trig_tails, 5)
 ```
+
+Calculate discounted probability mass at bigram lvl
+```{r}
+## Returns the total probability mass discounted from all observed bigrams
+## calculated from equation 14.  This is the amount of probability mass which
+## is redistributed to UNOBSERVED bigrams. If no bigrams starting with
+## unigram$ngram[1] exist, 0 is returned.
+##
+## unigram - single row, 2 column frequency table. The first column: ngram,
+##           contains the w_i-1 unigram (2nd word of the bigram prefix). The
+##           second column: freq, contains the frequency/count of this unigram.
+## bigrams - 2 column data.frame or data.table. The first column: ngram,
+##           contains all the bigrams in the corpus. The second column:
+##           freq, contains the frequency or count of each bigram.
+## bigDisc - amount to discount observed bigrams
+getAlphaBigram <- function(unigram, bigrams, bigDisc=0.5) {
+    # get all bigrams that start with unigram
+    regex <- sprintf("%s%s%s", "^", unigram$ngram[1], "_")
+    bigsThatStartWithUnig <- bigrams[grep(regex, bigrams$ngram),]
+    if(nrow(bigsThatStartWithUnig) < 1) return(0)
+    alphaBi <- 1 - (sum(bigsThatStartWithUnig$freq - bigDisc) / unigram$freq)
+
+    return(alphaBi)
+}
+
+unig <- str_split(bigPre, "_")[[1]][2]
+unig <- unigs[unigs$ngram == unig,]
+alpha_big <- getAlphaBigram(unig, bigrs, gamma2)
+alpha_big
+```
+
+Calculate backed-off probabilities q_bo for bigrams
+```{r}
+## Returns a character vector of backed off bigrams of the form w2_w1. These 
+## are all the (w_i-1, w) bigrams where w_i-1 is the tail word of the bigram
+## prefix bigPre and w are the tail words of unobserved bigrams that start with
+## w_i-1.
+##
+## bigPre - single-element char array of the form w2_w1 which are first two
+##          words of the trigram we are predicting the tail word of
+## unobsTrigTails - character vector that are tail words of unobserved trigrams
+getBoBigrams <- function(bigPre, unobsTrigTails) {
+    w_i_minus1 <- str_split(bigPre, "_")[[1]][2]
+    boBigrams <- paste(w_i_minus1, unobsTrigTails, sep = "_")
+    return(boBigrams)
+}
+
+## Returns a two column data.frame of backed-off bigrams in the first column
+## named ngram and their frequency/counts in the second column named freq.
+## 
+## bigPre -  single-element char array of the form w2_w1 which are first two
+##           words of the trigram we are predicting the tail word of
+## unobsTrigTails - character vector that are tail words of unobserved trigrams
+## bigrs - 2 column data.frame or data.table. The first column: ngram,
+##         contains all the bigrams in the corpus. The second column:
+##         freq, contains the frequency/count of each bigram.
+getObsBoBigrams <- function(bigPre, unobsTrigTails, bigrs) {
+    boBigrams <- getBoBigrams(bigPre, unobsTrigTails)
+    obs_bo_bigrams <- bigrs[bigrs$ngram %in% boBigrams, ]
+    return(obs_bo_bigrams)
+}
+
+## Returns a character vector of backed-off bigrams which are unobserved.
+##
+## bigPre -  single-element char array of the form w2_w1 which are first two
+##           words of the trigram we are predicting the tail word of
+## unobsTrigTails - character vector that are tail words of unobserved trigrams
+## obsBoBigram - data.frame which contains the observed bigrams in a column
+##               named ngram
+getUnobsBoBigrams <- function(bigPre, unobsTrigTails, obsBoBigram) {
+    boBigrams <- getBoBigrams(bigPre, unobsTrigTails)
+    unobs_bigs <- boBigrams[!(boBigrams %in% obsBoBigram$ngram)]
+    return(unobs_bigs)
+}
+
+## Returns a dataframe of 2 columns: ngram and probs.  Values in the ngram
+## column are bigrams of the form: w2_w1 which are observed as the last
+## two words in unobserved trigrams.  The values in the prob column are
+## q_bo(w1 | w2) calculated from from equation 10.
+##
+## obsBoBigrams - a dataframe with 2 columns: ngram and freq. The ngram column
+##                contains bigrams of the form w1_w2 which are observed bigrams
+##                that are the last 2 words of unobserved trigrams (i.e. "backed
+##                off" bigrams). The freq column contains integers that are
+##                the counts of these observed bigrams in the corpus.
+## unigs - 2 column data.frame of all the unigrams in the corpus:
+##         ngram = unigram
+##         freq = frequency/count of each unigram
+## bigDisc - amount to discount observed bigrams
+getObsBigProbs <- function(obsBoBigrams, unigs, bigDisc=0.5) {
+    first_words <- str_split_fixed(obsBoBigrams$ngram, "_", 2)[, 1]
+    first_word_freqs <- unigs[unigs$ngram %in% first_words, ]
+    obsBigProbs <- (obsBoBigrams$freq - bigDisc) / first_word_freqs$freq
+    obsBigProbs <- data.frame(ngram=obsBoBigrams$ngram, prob=obsBigProbs)
+
+    return(obsBigProbs)
+}
+
+## Returns a dataframe of 2 columns: ngram and prob.  Values in the ngram
+## column are unobserved bigrams of the form: w2_w1.  The values in the prob
+## column are the backed off probability estimates q_bo(w1 | w2) calculated
+## from from equation 16.
+##
+## unobsBoBigrams - character vector of unobserved backed off bigrams
+## unigs - 2 column data.frame of all the unigrams in the corpus:
+##         ngram = unigram
+##         freq = frequency/count of each unigram
+## alphaBig - total discounted probability mass at the bigram level
+getQboUnobsBigrams <- function(unobsBoBigrams, unigs, alphaBig) {
+    # get the unobserved bigram tails
+    qboUnobsBigs <- str_split_fixed(unobsBoBigrams, "_", 2)[, 2]
+    w_in_Aw_iminus1 <- unigs[!(unigs$ngram %in% qboUnobsBigs), ]
+    # convert to data.frame with counts
+    qboUnobsBigs <- unigs[unigs$ngram %in% qboUnobsBigs, ]
+    denom <- sum(qboUnobsBigs$freq)
+    # converts counts to probabilities
+    qboUnobsBigs <- data.frame(ngram=unobsBoBigrams,
+                               prob=(alphaBig * qboUnobsBigs$freq / denom))
+
+    return(qboUnobsBigs)
+}
+
+bo_bigrams <- getBoBigrams(bigPre, unobs_trig_tails)  # get backed off bigrams
+# separate bigrams which use eqn 10 and those that use 16
+obs_bo_bigrams <- getObsBoBigrams(bigPre, unobs_trig_tails, bigrs)
+unobs_bo_bigrams <- getUnobsBoBigrams(bigPre, unobs_trig_tails, obs_bo_bigrams)
+# unobs_bo_bigrams = c("the_buy", "the_EOS", "the_paint", "the_sell", "the_the")
+# calc obs'd bigram prob's from eqn 10
+qbo_obs_bigrams <- getObsBigProbs(obs_bo_bigrams, unigs, gamma2) #ngram     probs
+# calc alpha_big & unobs'd bigram prob's from eqn 16             #the_house 0.3125
+unig <- str_split(bigPre, "_")[[1]][2]
+unig <- unigs[unigs$ngram == unig,]
+# distrib discounted bigram prob mass to unobs bigrams in prop to unigram ML
+qbo_unobs_bigrams <- getQboUnobsBigrams(unobs_bo_bigrams, unigs, alpha_big)
+qbo_bigrams <- rbind(qbo_obs_bigrams, qbo_unobs_bigrams)
+qbo_bigrams
+```
+
+Checking Bigram Calculations
+```{r}
+unobs <- qbo_bigrams[-1,]
+sum(unobs$prob)
+```
+
+Discounted probability mass in trigram level
+```{r}
+## Returns the total probability mass discounted from all observed trigrams.
+## calculated from equation 14. This is the amount of probability mass which is
+## redistributed to UNOBSERVED trigrams. If no trigrams starting with
+## bigram$ngram[1] exist, 1 is returned.
+##
+## obsTrigs - 2 column data.frame or data.table. The first column: ngram,
+##            contains all the observed trigrams that start with the bigram
+##            prefix we are attempting to the predict the next word of. The 
+##            second column: freq, contains the frequency/count of each trigram.
+## bigram - single row frequency table where the first col: ngram, is the bigram
+##          which are the first two words of unobserved trigrams we want to
+##          estimate probabilities of (same as bigPre in other functions listed
+##          prior) delimited with '_'. The second column: freq, is the
+##          frequency/count of the bigram listed in the ngram column.
+## triDisc - amount to discount observed trigrams
+getAlphaTrigram <- function(obsTrigs, bigram, triDisc=0.5) {
+    if(nrow(obsTrigs) < 1) return(1)
+    alphaTri <- 1 - sum((obsTrigs$freq - triDisc) / bigram$freq[1])
+
+    return(alphaTri)
+}
+
+bigram <- bigrs[bigrs$ngram %in% bigPre, ]
+alpha_trig <- getAlphaTrigram(obs_trigs, bigram, gamma3)
+alpha_trig
+```
+
+Calculate UNOBSERVED trigram probabilities
+```{r}
+## Returns a dataframe of 2 columns: ngram and prob.  Values in the ngram
+## column are unobserved trigrams of the form: w3_w2_w1.  The values in the prob
+## column are q_bo(w1 | w3, w2) calculated from equation 17.
+##
+## bigPre -  single-element char array of the form w2_w1 which are first two
+##           words of the trigram we are predicting the tail word of
+## qboObsBigrams - 2 column data.frame with the following columns -
+##                 ngram: observed bigrams of the form w2_w1
+##                 probs: the probability estimate for observed bigrams:
+##                        qbo(w1 | w2) calc'd from equation 10.
+## qboUnobsBigrams - 2 column data.frame with the following columns -
+##                   ngram: unobserved bigrams of the form w2_w1
+##                   probs: the probability estimate for unobserved bigrams
+##                          qbo(w1 | w2) calc'd from equation 16.
+## alphaTrig - total discounted probability mass at the trigram level
+getUnobsTriProbs <- function(bigPre, qboObsBigrams,
+                             qboUnobsBigrams, alphaTrig) {
+    qboBigrams <- rbind(qboObsBigrams, qboUnobsBigrams)
+    qboBigrams <- qboBigrams[order(-qboBigrams$prob), ]
+    sumQboBigs <- sum(qboBigrams$prob)
+    first_bigPre_word <- str_split(bigPre, "_")[[1]][1]
+    unobsTrigNgrams <- paste(first_bigPre_word, qboBigrams$ngram, sep="_")
+    unobsTrigProbs <- alphaTrig * qboBigrams$prob / sumQboBigs
+    unobsTrigDf <- data.frame(ngram=unobsTrigNgrams, prob=unobsTrigProbs)
+
+    return(unobsTrigDf)
+}
+
+qbo_unobs_trigrams <- getUnobsTriProbs(bigPre, qbo_obs_bigrams,
+                                       qbo_unobs_bigrams, alpha_trig)
+qbo_unobs_trigrams
+```
+
+Get prediction with the highest prob
+```{r}
+getPredictionMsg <- function(qbo_trigs) {
+    # pull off tail word of highest prob trigram
+    prediction <- str_split(qbo_trigs$ngram[1], "_")[[1]][3]
+    result <- sprintf("%s%s%s%.4f", "highest prob prediction is >>> ", prediction,
+                      " <<< which has probability = ", qbo_trigs$prob[1])
+    return(result)
+}
+
+qbo_trigrams <- rbind(qbo_obs_trigrams, qbo_unobs_trigrams)
+qbo_trigrams <- qbo_trigrams[order(-qbo_trigrams$prob), ]  # sort by desc prob
+out_msg <- getPredictionMsg(qbo_trigrams)
+out_msg
+```
+
+testing
+```{r}
+gamma2=0.7; gamma3=0.7  # initialize new discount rates
+bigPre = "is_my"
+
+obs_trigs <- getObsTrigs(bigPre, trigs)
+unobs_trig_tails <- getUnobsTrigTails(obs_trigs$ngram, unigs)
+bo_bigrams <- getBoBigrams(bigPre, unobs_trig_tails)
+# separate bigrams which use eqn 10 and those that use 16
+obs_bo_bigrams <- getObsBoBigrams(bigPre, unobs_trig_tails, bigrs)
+unobs_bo_bigrams <- getUnobsBoBigrams(bigPre, unobs_trig_tails, obs_bo_bigrams)
+# calc obs'd bigram prob's from eqn 10
+qbo_obs_bigrams <- getObsBigProbs(obs_bo_bigrams, unigs, gamma2)
+# calc alpha_big & unobs'd bigram prob's from eqn 16
+unig <- str_split(bigPre, "_")[[1]][2]
+unig <- unigs[unigs$ngram == unig,]
+alpha_big <- getAlphaBigram(unig, bigrs, gamma2)
+# distrib discounted bigram prob mass to unobs bigrams in prop to unigram ML
+qbo_unobs_bigrams <- getQboUnobsBigrams(unobs_bo_bigrams, unigs, alpha_big)
+# calc trigram probabilities - start with observed trigrams: eqn 12
+qbo_obs_trigrams <- getObsTriProbs(obs_trigs, bigrs, bigPre, gamma3)
+# finally, calc trigram unobserved probabilities: eqn 17
+bigram <- bigrs[bigrs$ngram %in% bigPre, ]
+alpha_trig <- getAlphaTrigram(obs_trigs, bigram, gamma3)
+qbo_unobs_trigrams <- getUnobsTriProbs(bigPre, qbo_obs_bigrams,
+                                       qbo_unobs_bigrams, alpha_trig)
+qbo_trigrams <- rbind(qbo_obs_trigrams, qbo_unobs_trigrams)
+qbo_trigrams <- qbo_trigrams[order(-qbo_trigrams$prob), ]
+getPredictionMsg(qbo_trigrams)
+```

### 2019-11-12 temp.R

@@ -1,0 +1,32 @@
+con <- file("./en_US.blogs.txt", "r") 
+blog = readLines(con, skipNul = T)
+close(con)
+
+#takin 15% of data for memory reasons
+set.seed(12345)
+x = sample(899288, 1350, replace = F)
+
+train = blog[x]
+
+BG <- corpus(train)
+rm(con, train, blog, x)
+summary(BG, 5)
+
+
+## usual suspects
+getFreqs = function(dat, ng) {
+        dat.dfm = dfm(dat, ngrams = ng, remove_punct = T, remove_numbers = T,
+                      remove = stopwords("english"))
+        dat.freq = docfreq(dat.dfm)
+        dat.freq = dat.freq[sort(names(dat.freq))] 
+        return(dat.freq)
+}
+
+
+getTables = function(dat, ng) {
+        ngrams = getFreqs(dat = dat, ng = ng)
+        ngrams_dt = data.table(ngram = names(ngrams), freq = ngrams)
+        return(ngrams_dt)
+}
+
+

### 2019-11-13 draft1.Rmd
papa loves mambo

@@
 gamma3 <- 0.5  # trigram discount
-bigPre <- 'is_book'
+bigPre <- 'and_tell'
 
@@
 ##           two words of the trigram we are predicting the tail word of
-## trigrams - 2 column data.frame or data.table. The first column: ngram,
+## trigrams - 2 column data.table. The first column: ngram,
 ##            contains all the trigrams in the corpus. The second column:
 ##            freq, contains the frequency/count of each trigram.
+getODTrigs = function(bigPre, trigrams) {
+    tri.obs = data.frame(ngrams = vector(mode = 'character', length = 0),
+                          freq = vector(mode = 'integer', length = 0))
+    regex = sprintf("%s%s%s", "^", bigPre, "_")
+    indices = grep(regex, trigrams$ngram)
+    if(length(indices > 0)) {
+        tri.obs = trigrams[indices, ]
+    }
+    return(tri.obs)
+}
+
+#mikes's code
 getObsTrigs <- function(bigPre, trigrams) {
@@
 ## triDisc - amount to discount observed trigrams
+getODTProbs = function(obsTrigs, bigrs, bigPre, triDisc = 0.5) {
+    #return NULL if no trigrams starting with bigPre are observed
+    if(nrow(obsTrigs) <1) return(NULL)
+    #get total count of the observed trigrams
+    obsCount = filter(bigrs, ngram == bigPre)$freq
+    obsTrigProbs = mutate(obsTrigs, freq = ((freq - triDisc)/obsCount))
+    colnames(obsTrigProbs) = c("ngram", "prob")
+
+    return(obsTrigProbs)
+}
+
+#Mike's code
 getObsTriProbs <- function(obsTrigs, bigrs, bigPre, triDisc=0.5) {
@@
 ##         freq = frequency/count of each unigram
+getUODTTails = function(obsTrigs, unigs) {
+
+}
+
+#mike's code
 getUnobsTrigTails <- function(obsTrigs, unigs) {

### 2019-11-15 draft1.Rmd
some progress

@@
 gamma3 <- 0.5  # trigram discount
-bigPre <- 'and_tell'
+bigPre <- 'is_book'
 
@@
 
-obs_trigs <- getObsTrigs(bigPre, trigs)  # get trigrams and counts
+obs_trigs <- getODTrigs(bigPre, trigs)  # get trigrams and counts
 # convert counts to probabilities
-qbo_obs_trigrams <- getObsTriProbs(obs_trigs, bigrs, bigPre, gamma3)
+qbo_obs_trigrams <- getODTProbs(obs_trigs, bigrs, bigPre, gamma3)
 qbo_obs_trigrams
@@
 getUODTTails = function(obsTrigs, unigs) {
-
+    tails = str_split_fixed(obsTrigs, "_", 3)[, 3]
+    unobs_tails = unigs[!(unigs$ngram %in% tails), ]$ngram
+    return(unobs_tails)
 }
@@
 
-unobs_trig_tails <- getUnobsTrigTails(obs_trigs$ngram, unigs)
-head(unobs_trig_tails, 5)
+unobs_trig_tails <- getUODTTails(obs_trigs$ngram, unigs)
+head(unobs_trig_tails, 10)
 ```
@@
 ## bigDisc - amount to discount observed bigrams
+getAlphaBi = function(unigram, bigrams, bigDisc = 0.5) {
+    regex = sprintf("%s%s%s", "^", unigram$ngram[1], "_")
+    bgrs_with_unig = bigrams[grep(regex, bigrams$ngram), ]
+    if(nrow(bgrs_with_unig) < 1) return(0)
+    alphaBi = 1 - (sum(bgrs_with_unig$freq - bigDisc) / unigram$freq)
+    return(alphaBi)
+}
+#returns 0 if no observed bigrams
+
+#mike's code
 getAlphaBigram <- function(unigram, bigrams, bigDisc=0.5) {
@@
 unig <- unigs[unigs$ngram == unig,]
-alpha_big <- getAlphaBigram(unig, bigrs, gamma2)
+alpha_big <- getAlphaBi(unig, bigrs, gamma2)
 alpha_big
@@
 ```{r}
-## Returns a character vector of backed off bigrams of the form w2_w1. These 
-## are all the (w_i-1, w) bigrams where w_i-1 is the tail word of the bigram
+## Returns a character vector of backed off bigrams of the form w2_w1. 
+## These are all the (w_i-1, w) bigrams where w_i-1 is the tail word of the bigram
 ## prefix bigPre and w are the tail words of unobserved bigrams that start with
@@
 ## unobsTrigTails - character vector that are tail words of unobserved trigrams
+getBOBigrs = function(bigPre, UODTrigTails) {
+    w_i_1 = str_split(bigPre, "_")[[1]][2]
+    boBigrams = paste(w_i_1, UODTrigTails, sep = "_")
+    return(boBigrams)
+}
+
+#mike's code
 getBoBigrams <- function(bigPre, unobsTrigTails) {
@@
 ##         freq, contains the frequency/count of each bigram.
+getODBOBigrs = function(bigPre, UODTrigTails, bigrs) {
+    boBigrams = getBOBigrs(bigPre, UODTrigTails)
+    obs_bo_bigrs = bigrs[bigrs$ngram %in% boBigrams, ]
+    return(obs_bo_bigrs)
+}
+
+#mike's code
 getObsBoBigrams <- function(bigPre, unobsTrigTails, bigrs) {
@@
 ##               named ngram
+getUODBOBigrs = function(bigPre, UODTrigTails, ODBOBigrs) {
+    boBigrs = getBOBigrs(bigPre, UODTrigTails)
+    unobs_bigrs = boBigrs[!(boBigrs %in% ODBOBigrs$ngram)]
+    return(unobs_bigrs)
+}
+
+#mike's code
 getUnobsBoBigrams <- function(bigPre, unobsTrigTails, obsBoBigram) {
@@
 ## bigDisc - amount to discount observed bigrams
+getODBigrsProbs = function(ODBOBigrs, unigs, bigDisc = 0.5) {
+    first_words = str_split_fixed(ODBOBigrs$ngram, "_", 2)[ ,1]
+    first_word_freqs = unigs[unigs$ngram %in% first_words, ]
+    ODBigrsProbs = (ODBOBigrs$freq - bigDisc) / first_word_freqs$freq
+    ODBigrsProbs = data.frame(ngram = ODBOBigrs$ngram, prob = ODBigrsProbs)
+
+    return(ODBigrsProbs)
+}
+
+#mike's code
 getObsBigProbs <- function(obsBoBigrams, unigs, bigDisc=0.5) {
@@
 ## alphaBig - total discounted probability mass at the bigram level
+# CHECK
+getUODBOBigrsProbs = function(UODBOBigrs, unigs, alphaBig) {
+    #unobserved bigram tails
+    UODBigrsProbs = str_split_fixed(UODBOBigrs, "_", 2)[, 2]
+    w_Aw_i_1 = unigs[!(unigs$ngram %in% UODBigrsProbs), ]
+    #dataframe with counts
+    UODBigrsProbs = unigs[unigs$ngram %in% UODBigrsProbs, ]
+    qD = sum(UODBigrsProbs$freq)
+    #probabilities
+    UODBigrsProbs = data.frame(ngram = UODBOBigrs, 
+                               prob = (alphaBig * UODBigrsProbs$freq / qD))
+    return(UODBigrsProbs)
+}
+
+#mike's code
 getQboUnobsBigrams <- function(unobsBoBigrams, unigs, alphaBig) {
@@
     # convert to data.frame with counts
+    # PROBLEM HERE, w_in_Aw_iminus1 is not used to get qboUnobsBigs
     qboUnobsBigs <- unigs[unigs$ngram %in% qboUnobsBigs, ]
@@
 }
+
+#get backed off bigrams
+bo_bigrs = getBOBigrs(bigPre, unobs_trig_tails)
+#separate observed bigrams from unobserved bigrams
+obs_bo_bigrs = getODBOBigrs(bigPre, unobs_trig_tails, bigrs)
+unobs_bo_bigrs = getUODBOBigrs(bigPre, unobs_trig_tails, obs_bo_bigrs)
+#qbo
+qbo_obs_bigrams = getODBigrsProbs(obs_bo_bigrs, unigs, gamma2)
+
+unig <- str_split(bigPre, "_")[[1]][2]
+unig <- unigs[unigs$ngram == unig,]
+#qbo redistributed
+qbo_unobs_bigrams = getUODBOBigrsProbs(unobs_bo_bigrs, unigs, alpha_big)
 

### 2019-11-15 temp.R

@@ -29,4 +29,15 @@ getTables = function(dat, ng) {
         return(ngrams_dt)
 }
 
+# UODBOBigrs = unobs_bo_bigrs, unigs, alphaBig = alpha_big
+
+UODBigrsProbs = str_split_fixed(UODBOBigrs, "_", 2)[, 2]
+w_Aw_i_1 = unigs[!(unigs$ngram %in% UODBigrsProbs), ]
+#dataframe with counts
+UODBigrsProbs = unigs[unigs$ngram %in% w_Aw_i_1, ]
+qD = sum(UODBigrsProbs$freq)
+#probabilities
+UODBigrsProbs = data.frame(ngram = UODBOBigrs, 
+                           prob = (alphaBig * UODBigrsProbs$freq / qD))
+
 

### 2019-11-16 Notes.Rmd
stupid backoff

@@
 ```
+
+For example, given input of "another brick in the" one would start with a 5-gram model, use the first four words as the base value, and the fifth word as the predicted value.
+The correct prediction is "wall".
+Rendered in a very simple Shiny app, it looks like this:
+
+If you can't find a match with the 4-gram base and a prediction, then you take the three rightmost words "brick in the" to predict the last word in a 4-gram.
+If that fails, you take "in the" to predict the last word in a 3-gram, and so on until you decide how to handle the situation where you can't find a valid prediction from the word "the".
 

### 2019-11-16 al_stup.Rmd
stupid backoff

--- 2019-11-13 draft1.Rmd
@@
 ---
-title: "SwiftKey Capstone Proj. Milestone Report"
+title: "stupid backoff"
 author: "Yigit Ozan Berk"
-date: "11/9/2019"
+date: "11/16/2019"
 output: html_document
@@
 
-# Initiation
 ```{r}
@@
 
+# Algorithm
+
 Probs of words completing OBSERVED trigrams
 ```{r}
-gamma2 <- 0.5  # bigram discount
-gamma3 <- 0.5  # trigram discount
-bigPre <- 'and_tell'
+bigPre <- 'is_book'
 
@@
 
-#mikes's code
-getObsTrigs <- function(bigPre, trigrams) {
-    trigs.winA <- data.frame(ngrams=vector(mode = 'character', length = 0),
-                             freq=vector(mode = 'integer', length = 0))
-    regex <- sprintf("%s%s%s", "^", bigPre, "_")
-    trigram_indices <- grep(regex, trigrams$ngram)
-    if(length(trigram_indices) > 0) {
-        trigs.winA <- trigrams[trigram_indices, ]
+getODBigs = function(bigPre, bigrams) {
+       big.obs = data.frame(ngrams = vector(mode = 'character', length = 0),
+                            freq = vector(mode = 'integer', length = 0))
+       w_i_1 <- str_split(bigPre, "_")[[1]][2]
+       regex = sprintf("%s%s%s", "^", w_i_1, "_")
+       indices = grep(regex, bigrams$ngram)
+       if(length(indices > 0)) {
+               big.obs = bigrams[indices, ]
+       }
+       return(big.obs)
 }
 
-    return(trigs.winA)
-}
-
-## Returns a two column data.frame of observed trigrams that start with bigram
-## prefix bigPre in the first column named ngram and the probabilities
-## q_bo(w_i | w_i-2, w_i-1) in the second column named prob calculated from
-## eqn 12. If no observed trigrams starting with bigPre exist, NULL is returned.
-##
-## obsTrigs - 2 column data.frame or data.table. The first column: ngram,
-##            contains all the observed trigrams that start with the bigram
-##            prefix bigPre which we are attempting to the predict the next
-##            word of in a give phrase. The second column: freq, contains the
-##            frequency/count of each trigram.
-## bigrs - 2 column data.frame or data.table. The first column: ngram,
-##         contains all the bigrams in the corpus. The second column:
-##         freq, contains the frequency/count of each bigram.
-## bigPre -  single-element char array of the form w2_w1 which are first two
-##           words of the trigram we are predicting the tail word of
-## triDisc - amount to discount observed trigrams
-getODTProbs = function(obsTrigs, bigrs, bigPre, triDisc = 0.5) {
-    #return NULL if no trigrams starting with bigPre are observed
-    if(nrow(obsTrigs) <1) return(NULL)
-    #get total count of the observed trigrams
-    obsCount = filter(bigrs, ngram == bigPre)$freq
-    obsTrigProbs = mutate(obsTrigs, freq = ((freq - triDisc)/obsCount))
-    colnames(obsTrigProbs) = c("ngram", "prob")
-
-    return(obsTrigProbs)
-}
-
-#Mike's code
-getObsTriProbs <- function(obsTrigs, bigrs, bigPre, triDisc=0.5) {
-    if(nrow(obsTrigs) < 1) return(NULL)
-    obsCount <- filter(bigrs, ngram==bigPre)$freq[1]
-    obsTrigProbs <- mutate(obsTrigs, freq=((freq - triDisc) / obsCount))
-    colnames(obsTrigProbs) <- c("ngram", "prob")
-
-    return(obsTrigProbs)
-}
-
-obs_trigs <- getObsTrigs(bigPre, trigs)  # get trigrams and counts
-# convert counts to probabilities
-qbo_obs_trigrams <- getObsTriProbs(obs_trigs, bigrs, bigPre, gamma3)
-qbo_obs_trigrams
-```
-
-Find UNOBSERVED trigram tail words
-```{r}
-## Returns a character vector which are the tail words of unobserved trigrams
-## that start with the first two words of obsTrigs (aka the bigram prefix).
-## These are the words w in the set B(w_i-2, w_i-1) as defined in the section
-## describing the details of equation 17.
-##
-## obsTrigs - character vector of observed trigrams delimited by _ of the form:
-##            w3_w2_w1 where w3_w2 is the bigram prefix
-## unigs - 2 column data.frame of all the unigrams in the corpus:
-##         ngram = unigram
-##         freq = frequency/count of each unigram
-getUODTTails = function(obsTrigs, unigs) {
+getODUnigs = function(bigPre, unigrams) {
 
@@
 
-#mike's code
-getUnobsTrigTails <- function(obsTrigs, unigs) {
-    obs_trig_tails <- str_split_fixed(obsTrigs, "_", 3)[, 3]
-    unobs_trig_tails <- unigs[!(unigs$ngram %in% obs_trig_tails), ]$ngram
-    return(unobs_trig_tails)
+getTrigProbs = function(bigPre, trigrams, bigrams) {
+
 }
 
-unobs_trig_tails <- getUnobsTrigTails(obs_trigs$ngram, unigs)
-head(unobs_trig_tails, 5)
+S3 = getODTrigs(bigPre, trigs) # get observed trigrams and counts
+S2 = getODBigs(bigPre, bigrs) # get observed bigrams and counts
+
 ```
-
-Calculate discounted probability mass at bigram lvl
-```{r}
-## Returns the total probability mass discounted from all observed bigrams
-## calculated from equation 14.  This is the amount of probability mass which
-## is redistributed to UNOBSERVED bigrams. If no bigrams starting with
-## unigram$ngram[1] exist, 0 is returned.
-##
-## unigram - single row, 2 column frequency table. The first column: ngram,
-##           contains the w_i-1 unigram (2nd word of the bigram prefix). The
-##           second column: freq, contains the frequency/count of this unigram.
-## bigrams - 2 column data.frame or data.table. The first column: ngram,
-##           contains all the bigrams in the corpus. The second column:
-##           freq, contains the frequency or count of each bigram.
-## bigDisc - amount to discount observed bigrams
-getAlphaBigram <- function(unigram, bigrams, bigDisc=0.5) {
-    # get all bigrams that start with unigram
-    regex <- sprintf("%s%s%s", "^", unigram$ngram[1], "_")
-    bigsThatStartWithUnig <- bigrams[grep(regex, bigrams$ngram),]
-    if(nrow(bigsThatStartWithUnig) < 1) return(0)
-    alphaBi <- 1 - (sum(bigsThatStartWithUnig$freq - bigDisc) / unigram$freq)
-
-    return(alphaBi)
-}
-
-unig <- str_split(bigPre, "_")[[1]][2]
-unig <- unigs[unigs$ngram == unig,]
-alpha_big <- getAlphaBigram(unig, bigrs, gamma2)
-alpha_big
-```
-
-Calculate backed-off probabilities q_bo for bigrams
-```{r}
-## Returns a character vector of backed off bigrams of the form w2_w1. These 
-## are all the (w_i-1, w) bigrams where w_i-1 is the tail word of the bigram
-## prefix bigPre and w are the tail words of unobserved bigrams that start with
-## w_i-1.
-##
-## bigPre - single-element char array of the form w2_w1 which are first two
-##          words of the trigram we are predicting the tail word of
-## unobsTrigTails - character vector that are tail words of unobserved trigrams
-getBoBigrams <- function(bigPre, unobsTrigTails) {
-    w_i_minus1 <- str_split(bigPre, "_")[[1]][2]
-    boBigrams <- paste(w_i_minus1, unobsTrigTails, sep = "_")
-    return(boBigrams)
-}
-
-## Returns a two column data.frame of backed-off bigrams in the first column
-## named ngram and their frequency/counts in the second column named freq.
-## 
-## bigPre -  single-element char array of the form w2_w1 which are first two
-##           words of the trigram we are predicting the tail word of
-## unobsTrigTails - character vector that are tail words of unobserved trigrams
-## bigrs - 2 column data.frame or data.table. The first column: ngram,
-##         contains all the bigrams in the corpus. The second column:
-##         freq, contains the frequency/count of each bigram.
-getObsBoBigrams <- function(bigPre, unobsTrigTails, bigrs) {
-    boBigrams <- getBoBigrams(bigPre, unobsTrigTails)
-    obs_bo_bigrams <- bigrs[bigrs$ngram %in% boBigrams, ]
-    return(obs_bo_bigrams)
-}
-
-## Returns a character vector of backed-off bigrams which are unobserved.
-##
-## bigPre -  single-element char array of the form w2_w1 which are first two
-##           words of the trigram we are predicting the tail word of
-## unobsTrigTails - character vector that are tail words of unobserved trigrams
-## obsBoBigram - data.frame which contains the observed bigrams in a column
-##               named ngram
-getUnobsBoBigrams <- function(bigPre, unobsTrigTails, obsBoBigram) {
-    boBigrams <- getBoBigrams(bigPre, unobsTrigTails)
-    unobs_bigs <- boBigrams[!(boBigrams %in% obsBoBigram$ngram)]
-    return(unobs_bigs)
-}
-
-## Returns a dataframe of 2 columns: ngram and probs.  Values in the ngram
-## column are bigrams of the form: w2_w1 which are observed as the last
-## two words in unobserved trigrams.  The values in the prob column are
-## q_bo(w1 | w2) calculated from from equation 10.
-##
-## obsBoBigrams - a dataframe with 2 columns: ngram and freq. The ngram column
-##                contains bigrams of the form w1_w2 which are observed bigrams
-##                that are the last 2 words of unobserved trigrams (i.e. "backed
-##                off" bigrams). The freq column contains integers that are
-##                the counts of these observed bigrams in the corpus.
-## unigs - 2 column data.frame of all the unigrams in the corpus:
-##         ngram = unigram
-##         freq = frequency/count of each unigram
-## bigDisc - amount to discount observed bigrams
-getObsBigProbs <- function(obsBoBigrams, unigs, bigDisc=0.5) {
-    first_words <- str_split_fixed(obsBoBigrams$ngram, "_", 2)[, 1]
-    first_word_freqs <- unigs[unigs$ngram %in% first_words, ]
-    obsBigProbs <- (obsBoBigrams$freq - bigDisc) / first_word_freqs$freq
-    obsBigProbs <- data.frame(ngram=obsBoBigrams$ngram, prob=obsBigProbs)
-
-    return(obsBigProbs)
-}
-
-## Returns a dataframe of 2 columns: ngram and prob.  Values in the ngram
-## column are unobserved bigrams of the form: w2_w1.  The values in the prob
-## column are the backed off probability estimates q_bo(w1 | w2) calculated
-## from from equation 16.
-##
-## unobsBoBigrams - character vector of unobserved backed off bigrams
-## unigs - 2 column data.frame of all the unigrams in the corpus:
-##         ngram = unigram
-##         freq = frequency/count of each unigram
-## alphaBig - total discounted probability mass at the bigram level
-getQboUnobsBigrams <- function(unobsBoBigrams, unigs, alphaBig) {
-    # get the unobserved bigram tails
-    qboUnobsBigs <- str_split_fixed(unobsBoBigrams, "_", 2)[, 2]
-    w_in_Aw_iminus1 <- unigs[!(unigs$ngram %in% qboUnobsBigs), ]
-    # convert to data.frame with counts
-    qboUnobsBigs <- unigs[unigs$ngram %in% qboUnobsBigs, ]
-    denom <- sum(qboUnobsBigs$freq)
-    # converts counts to probabilities
-    qboUnobsBigs <- data.frame(ngram=unobsBoBigrams,
-                               prob=(alphaBig * qboUnobsBigs$freq / denom))
-
-    return(qboUnobsBigs)
-}
-
-bo_bigrams <- getBoBigrams(bigPre, unobs_trig_tails)  # get backed off bigrams
-# separate bigrams which use eqn 10 and those that use 16
-obs_bo_bigrams <- getObsBoBigrams(bigPre, unobs_trig_tails, bigrs)
-unobs_bo_bigrams <- getUnobsBoBigrams(bigPre, unobs_trig_tails, obs_bo_bigrams)
-# unobs_bo_bigrams = c("the_buy", "the_EOS", "the_paint", "the_sell", "the_the")
-# calc obs'd bigram prob's from eqn 10
-qbo_obs_bigrams <- getObsBigProbs(obs_bo_bigrams, unigs, gamma2) #ngram     probs
-# calc alpha_big & unobs'd bigram prob's from eqn 16             #the_house 0.3125
-unig <- str_split(bigPre, "_")[[1]][2]
-unig <- unigs[unigs$ngram == unig,]
-# distrib discounted bigram prob mass to unobs bigrams in prop to unigram ML
-qbo_unobs_bigrams <- getQboUnobsBigrams(unobs_bo_bigrams, unigs, alpha_big)
-qbo_bigrams <- rbind(qbo_obs_bigrams, qbo_unobs_bigrams)
-qbo_bigrams
-```
-
-Checking Bigram Calculations
-```{r}
-unobs <- qbo_bigrams[-1,]
-sum(unobs$prob)
-```
-
-Discounted probability mass in trigram level
-```{r}
-## Returns the total probability mass discounted from all observed trigrams.
-## calculated from equation 14. This is the amount of probability mass which is
-## redistributed to UNOBSERVED trigrams. If no trigrams starting with
-## bigram$ngram[1] exist, 1 is returned.
-##
-## obsTrigs - 2 column data.frame or data.table. The first column: ngram,
-##            contains all the observed trigrams that start with the bigram
-##            prefix we are attempting to the predict the next word of. The 
-##            second column: freq, contains the frequency/count of each trigram.
-## bigram - single row frequency table where the first col: ngram, is the bigram
-##          which are the first two words of unobserved trigrams we want to
-##          estimate probabilities of (same as bigPre in other functions listed
-##          prior) delimited with '_'. The second column: freq, is the
-##          frequency/count of the bigram listed in the ngram column.
-## triDisc - amount to discount observed trigrams
-getAlphaTrigram <- function(obsTrigs, bigram, triDisc=0.5) {
-    if(nrow(obsTrigs) < 1) return(1)
-    alphaTri <- 1 - sum((obsTrigs$freq - triDisc) / bigram$freq[1])
-
-    return(alphaTri)
-}
-
-bigram <- bigrs[bigrs$ngram %in% bigPre, ]
-alpha_trig <- getAlphaTrigram(obs_trigs, bigram, gamma3)
-alpha_trig
-```
-
-Calculate UNOBSERVED trigram probabilities
-```{r}
-## Returns a dataframe of 2 columns: ngram and prob.  Values in the ngram
-## column are unobserved trigrams of the form: w3_w2_w1.  The values in the prob
-## column are q_bo(w1 | w3, w2) calculated from equation 17.
-##
-## bigPre -  single-element char array of the form w2_w1 which are first two
-##           words of the trigram we are predicting the tail word of
-## qboObsBigrams - 2 column data.frame with the following columns -
-##                 ngram: observed bigrams of the form w2_w1
-##                 probs: the probability estimate for observed bigrams:
-##                        qbo(w1 | w2) calc'd from equation 10.
-## qboUnobsBigrams - 2 column data.frame with the following columns -
-##                   ngram: unobserved bigrams of the form w2_w1
-##                   probs: the probability estimate for unobserved bigrams
-##                          qbo(w1 | w2) calc'd from equation 16.
-## alphaTrig - total discounted probability mass at the trigram level
-getUnobsTriProbs <- function(bigPre, qboObsBigrams,
-                             qboUnobsBigrams, alphaTrig) {
-    qboBigrams <- rbind(qboObsBigrams, qboUnobsBigrams)
-    qboBigrams <- qboBigrams[order(-qboBigrams$prob), ]
-    sumQboBigs <- sum(qboBigrams$prob)
-    first_bigPre_word <- str_split(bigPre, "_")[[1]][1]
-    unobsTrigNgrams <- paste(first_bigPre_word, qboBigrams$ngram, sep="_")
-    unobsTrigProbs <- alphaTrig * qboBigrams$prob / sumQboBigs
-    unobsTrigDf <- data.frame(ngram=unobsTrigNgrams, prob=unobsTrigProbs)
-
-    return(unobsTrigDf)
-}
-
-qbo_unobs_trigrams <- getUnobsTriProbs(bigPre, qbo_obs_bigrams,
-                                       qbo_unobs_bigrams, alpha_trig)
-qbo_unobs_trigrams
-```
-
-Get prediction with the highest prob
-```{r}
-getPredictionMsg <- function(qbo_trigs) {
-    # pull off tail word of highest prob trigram
-    prediction <- str_split(qbo_trigs$ngram[1], "_")[[1]][3]
-    result <- sprintf("%s%s%s%.4f", "highest prob prediction is >>> ", prediction,
-                      " <<< which has probability = ", qbo_trigs$prob[1])
-    return(result)
-}
-
-qbo_trigrams <- rbind(qbo_obs_trigrams, qbo_unobs_trigrams)
-qbo_trigrams <- qbo_trigrams[order(-qbo_trigrams$prob), ]  # sort by desc prob
-out_msg <- getPredictionMsg(qbo_trigrams)
-out_msg
-```
-
-testing
-```{r}
-gamma2=0.7; gamma3=0.7  # initialize new discount rates
-bigPre = "is_my"
-
-obs_trigs <- getObsTrigs(bigPre, trigs)
-unobs_trig_tails <- getUnobsTrigTails(obs_trigs$ngram, unigs)
-bo_bigrams <- getBoBigrams(bigPre, unobs_trig_tails)
-# separate bigrams which use eqn 10 and those that use 16
-obs_bo_bigrams <- getObsBoBigrams(bigPre, unobs_trig_tails, bigrs)
-unobs_bo_bigrams <- getUnobsBoBigrams(bigPre, unobs_trig_tails, obs_bo_bigrams)
-# calc obs'd bigram prob's from eqn 10
-qbo_obs_bigrams <- getObsBigProbs(obs_bo_bigrams, unigs, gamma2)
-# calc alpha_big & unobs'd bigram prob's from eqn 16
-unig <- str_split(bigPre, "_")[[1]][2]
-unig <- unigs[unigs$ngram == unig,]
-alpha_big <- getAlphaBigram(unig, bigrs, gamma2)
-# distrib discounted bigram prob mass to unobs bigrams in prop to unigram ML
-qbo_unobs_bigrams <- getQboUnobsBigrams(unobs_bo_bigrams, unigs, alpha_big)
-# calc trigram probabilities - start with observed trigrams: eqn 12
-qbo_obs_trigrams <- getObsTriProbs(obs_trigs, bigrs, bigPre, gamma3)
-# finally, calc trigram unobserved probabilities: eqn 17
-bigram <- bigrs[bigrs$ngram %in% bigPre, ]
-alpha_trig <- getAlphaTrigram(obs_trigs, bigram, gamma3)
-qbo_unobs_trigrams <- getUnobsTriProbs(bigPre, qbo_obs_bigrams,
-                                       qbo_unobs_bigrams, alpha_trig)
-qbo_trigrams <- rbind(qbo_obs_trigrams, qbo_unobs_trigrams)
-qbo_trigrams <- qbo_trigrams[order(-qbo_trigrams$prob), ]
-getPredictionMsg(qbo_trigrams)
-```

### 2019-11-16 draft1.Rmd
stupid backoff

@@
 qbo_unobs_bigrams = getUODBOBigrsProbs(unobs_bo_bigrs, unigs, alpha_big)
-
-bo_bigrams <- getBoBigrams(bigPre, unobs_trig_tails)  # get backed off bigrams
-# separate bigrams which use eqn 10 and those that use 16
-obs_bo_bigrams <- getObsBoBigrams(bigPre, unobs_trig_tails, bigrs)
-unobs_bo_bigrams <- getUnobsBoBigrams(bigPre, unobs_trig_tails, obs_bo_bigrams)
-# unobs_bo_bigrams = c("the_buy", "the_EOS", "the_paint", "the_sell", "the_the")
-# calc obs'd bigram prob's from eqn 10
-qbo_obs_bigrams <- getObsBigProbs(obs_bo_bigrams, unigs, gamma2) #ngram     probs
-# calc alpha_big & unobs'd bigram prob's from eqn 16             #the_house 0.3125
-unig <- str_split(bigPre, "_")[[1]][2]
-unig <- unigs[unigs$ngram == unig,]
-# distrib discounted bigram prob mass to unobs bigrams in prop to unigram ML
-qbo_unobs_bigrams <- getQboUnobsBigrams(unobs_bo_bigrams, unigs, alpha_big)
-qbo_bigrams <- rbind(qbo_obs_bigrams, qbo_unobs_bigrams)
-qbo_bigrams
+qbo_bigrams = rbind(qbo_obs_bigrams, qbo_unobs_bigrams)
+head(qbo_bigrams)
 ```
@@
 ## triDisc - amount to discount observed trigrams
+getAlphaTri = function(obsTrigs, bigram, triDisc = 0.5) {
+    if(nrow(obsTrigs) < 1) return(1)
+    alphaTri = 1 - sum((obsTrigs$freq - triDisc) / bigram$freq[1])
+
+    return(alphaTri)
+}
+
+#mike's code
 getAlphaTrigram <- function(obsTrigs, bigram, triDisc=0.5) {
@@
 bigram <- bigrs[bigrs$ngram %in% bigPre, ]
-alpha_trig <- getAlphaTrigram(obs_trigs, bigram, gamma3)
+alpha_trig <- getAlphaTri(obs_trigs, bigram, gamma3)
 alpha_trig
@@
 ## alphaTrig - total discounted probability mass at the trigram level
+getUODTrigProbs = function()
+
+#mike's code
 getUnobsTriProbs <- function(bigPre, qboObsBigrams,

### 2019-11-16 temp.R

@@ -41,3 +41,38 @@ UODBigrsProbs = data.frame(ngram = UODBOBigrs,
                            prob = (alphaBig * UODBigrsProbs$freq / qD))
 
 
+
+
+
+# Profanity
+
+library(magrittr)
+
+samplefile <- function(filename, fraction) {
+        system(paste("perl -ne 'print if (rand() < ",
+                     fraction, ")'", filename), intern=TRUE)
+}
+
+tokenize <- function(v) {
+        # Add spaces before and after punctuation,
+        # remove repeat spaces, and split the strings
+        gsub("([^ ])([.?!&])", "\\1 \\2 ", v)   %>%
+                gsub(pattern=" +", replacement=" ")     %>%
+                strsplit(split=" ")
+}
+
+profanity_filter <- function(tv) {
+        # Takes a tokenized vector
+        bad_words <- paste("([Ff][Uu][Cc][Kk]",
+                           "[Dd][Aa][Mm][Nn]",
+                           "[Ss$][Hh][Ii][Tt]",
+                           "[Aa@][Ss$][Ss$]",
+                           "[Aa@][Ss$][Ss$][Hh][Oo][Ll][Ee]",
+                           "[Cc][Uu][Nn][Tt]",
+                           "[Nn][Ii][Gg][Gg][Ee][Rr])", sep="|")
+        
+        lapply(tv, function(x) gsub(bad_words, " ", x)) %>%
+                lapply(FUN=function(x) x[!x == " "])
+}
+
+

### 2019-11-18 al_stup.Rmd
full algorithm

@@
 getFreqs = function(dat, ng) {
-        dat.dfm = dfm(dat, ngrams = ng, remove_punct = T, remove_numbers = T,
-                      remove = stopwords("english"))
+        dat.dfm = dfm(dat, ngrams = ng, remove_punct = T, remove_numbers = T
+                      #,remove = stopwords("english")
+                      )
         dat.freq = docfreq(dat.dfm)
@@
 ```{r}
-unigs = getTables(BG, 1)
-bigrs = getTables(BG, 2)
-trigs = getTables(BG, 3)
+BG.unigs = getTables(BG, 1)
+BG.bigrs = getTables(BG, 2)
+BG.trigs = getTables(BG, 3)
+rm(BG)
+
+NS.unigs = getTables(NS, 1)
+NS.bigrs = getTables(NS, 2)
+NS.trigs = getTables(NS, 3)
+rm(NS)
+
+TW.unigs = getTables(TW, 1)
+TW.bigrs = getTables(TW, 2)
+TW.trigs = getTables(TW, 3)
+rm(TW)
+
 ```
@@
 
+## Returns a two column data.frame of observed bigrams that start with the
+## unigram prefix (second word of bigPre) in the first column named ngram and
+## frequencies/counts in the second column named freq. If no observed bigrams
+## that start with second word of bigPre exist, an empty data.frame is returned.
+##
+## bigPre -  single-element char array of the form w2_w1 which are the first 
+##           two words of the trigram we are predicting the tail word of
+## trigrams - 2 column data.table. The first column: ngram,
+##            contains all the trigrams in the corpus. The second column:
+##            freq, contains the frequency/count of each trigram.
 getODBigs = function(bigPre, bigrams) {
@@
 
-getODUnigs = function(bigPre, unigrams) {
+#x = getTrigProbs(bigPre, trigs, bigrs, unigs)
+
+getTrigProbs = function(bigPre, trigrams, bigrams, unigrams) {
+        obs_trigs = getODTrigs(bigPre, trigrams)
+        regex = sprintf("%s%s%s", "^", bigPre, "$")
+        bigr_count = bigrams[grep(regex, bigrams$ngram), ]$freq
+
+        if(nrow(obs_trigs) < 1) {
+                bigr_probs = getBigProbs(bigPre, bigrams, unigrams)
+                probs = data.frame(ngram = bigr_probs$ngram, 
+                                   prob = bigr_probs$prob * 0.4)
+                return(probs)
+        } else {
+                probs = obs_trigs %>% rename( prob = freq) %>%
+                    mutate(prob = prob / bigr_count) %>% 
+                    arrange(desc(prob), ngram)
+                return(probs)
+        }
 
@@
 
-getTrigProbs = function(bigPre, trigrams, bigrams) {
-
+getBigProbs = function(bigPre, bigrams, unigrams) {
+    obs_bigs = getODBigs(bigPre, bigrams)
+    w_i_1 <- str_split(bigPre, "_")[[1]][2]
+    regex = sprintf("%s%s%s", "^", w_i_1, "$")
+    unig_count = unigrams[grep(regex, unigrams$ngram), ]$freq
+    if(nrow(obs_bigs) < 1) {
+        probs = getUnigProbs(unigrams) %>% mutate(prob = prob * 0.4)
+        return(probs)
+    } else {
+        probs = obs_bigs %>% rename(prob = freq) %>%
+            mutate(prob = prob/unig_count) %>%
+            arrange(desc(prob), ngram)
+        return(probs)
+    }
 }
 
-S3 = getODTrigs(bigPre, trigs) # get observed trigrams and counts
-S2 = getODBigs(bigPre, bigrs) # get observed bigrams and counts
+getUnigProbs = function(unigrams) {
+    probs = unigrams %>% rename(prob = freq) %>%
+        mutate(prob = prob/sum(prob)) %>%
+        arrange(desc(prob), ngram)
+    return(probs)
+}
 

### 2019-11-19 al_stup.Rmd
data.table modifications

@@
 ```{r}
-getFreqs = function(dat, ng) {
-        dat.dfm = dfm(dat, ngrams = ng, remove_punct = T, remove_numbers = T
-                      #,remove = stopwords("english")
-                      )
-        dat.freq = docfreq(dat.dfm)
-        dat.freq = dat.freq[sort(names(dat.freq))] 
-        return(dat.freq)
-}
-
 getTables = function(dat, ng) {
-        ngrams = getFreqs(dat = dat, ng = ng)
-        ngrams_dt = data.table(ngram = names(ngrams), freq = ngrams)
-        return(ngrams_dt)
+    dat.dfm = dfm(dat, ngrams = ng, remove_punct = T, remove_numbers = T)
+    dat.tbl = data.table(ngram = featnames(dat.dfm), count = colSums(dat.dfm),
+                         key = "ngram")
+    dat.tbl = dat.tbl[order(ngram)]
+    return(dat.tbl)
 }

### 2019-11-20 al_stup.Rmd
data.table transformation MAIN CHANGE IN ALGORITHM

--- 2019-11-16 al_stup.Rmd
@@
 ```{r}
-getFreqs = function(dat, ng) {
-        dat.dfm = dfm(dat, ngrams = ng, remove_punct = T, remove_numbers = T,
-                      remove = stopwords("english"))
-        dat.freq = docfreq(dat.dfm)
-        dat.freq = dat.freq[sort(names(dat.freq))] 
-        return(dat.freq)
-}
-
 getTables = function(dat, ng) {
-        ngrams = getFreqs(dat = dat, ng = ng)
-        ngrams_dt = data.table(ngram = names(ngrams), freq = ngrams)
-        return(ngrams_dt)
+    dat.dfm = dfm(dat, ngrams = ng, remove_punct = T, remove_numbers = T)
+    dat.tbl = data.table(ngram = featnames(dat.dfm), freq = colSums(dat.dfm),
+                         key = "ngram")
+    dat.tbl = dat.tbl[order(ngram)]
+    return(dat.tbl)
 }
@@
 ```{r}
-unigs = getTables(BG, 1)
-bigrs = getTables(BG, 2)
-trigs = getTables(BG, 3)
+BG.unigs = getTables(BG, 1)
+BG.bigrs = getTables(BG, 2)
+BG.trigs = getTables(BG, 3)
+rm(BG)
+
+NS.unigs = getTables(NS, 1)
+NS.bigrs = getTables(NS, 2)
+NS.trigs = getTables(NS, 3)
+rm(NS)
+
+unigs = getTables(TW, 1)
+bigrs = getTables(TW, 2)
+trigs = getTables(TW, 3)
+rm(TW)
+
 ```
@@
 
-## Returns a two column data.frame of observed trigrams that start with the
+## Returns a two column data.table of observed trigrams that start with the
 ## bigram prefix (bigPre) in the first column named ngram and
 ## frequencies/counts in the second column named freq. If no observed trigrams
-## that start with bigPre exist, an empty data.frame is returned.
+## that start with bigPre exist, an empty data.table is returned.
 ##
@@
 getODTrigs = function(bigPre, trigrams) {
-    tri.obs = data.frame(ngrams = vector(mode = 'character', length = 0),
-                          freq = vector(mode = 'integer', length = 0))
     regex = sprintf("%s%s%s", "^", bigPre, "_")
-    indices = grep(regex, trigrams$ngram)
-    if(length(indices > 0)) {
-        tri.obs = trigrams[indices, ]
-    }
+    tri.obs = trigrams[ngram %like% regex]
     return(tri.obs)
@@
 
+## Returns a two column data.table of observed bigrams that start with the
+## unigram prefix (second word of bigPre) in the first column named ngram and
+## frequencies/counts in the second column named freq. If no observed bigrams
+## that start with second word of bigPre exist, an empty data.table is returned.
+##
+## bigPre -  single-element char array of the form w2_w1 which are the first 
+##           two words of the trigram we are predicting the tail word of
+## trigrams - 2 column data.table. The first column: ngram,
+##            contains all the trigrams in the corpus. The second column:
+##            freq, contains the frequency/count of each trigram.
 getODBigs = function(bigPre, bigrams) {
-       big.obs = data.frame(ngrams = vector(mode = 'character', length = 0),
-                            freq = vector(mode = 'integer', length = 0))
-       w_i_1 <- str_split(bigPre, "_")[[1]][2]
+    w_i_1 = str_split(bigPre, "_")[[1]][2]
     regex = sprintf("%s%s%s", "^", w_i_1, "_")
-       indices = grep(regex, bigrams$ngram)
-       if(length(indices > 0)) {
-               big.obs = bigrams[indices, ]
-       }
+    big.obs = bigrams[ngram %like% regex]
     return(big.obs)
@@
 
-getODUnigs = function(bigPre, unigrams) {
+#x = getTrigProbs(bigPre, trigs, bigrs, unigs)
+
+getTrigProbs = function(bigPre, trigrams, bigrams, unigrams) {
+        obs_trigs = getODTrigs(bigPre, trigrams)
+        regex = sprintf("%s%s%s", "^", bigPre, "$")
+        #bigr_count = bigrams[grep(regex, bigrams$ngram), ]$freq
+        bigr_count = bigrams[ngram %like% regex]$freq
+
+        if(nrow(obs_trigs) < 1) {
+                bigr_probs = getBigProbs(bigPre, bigrams, unigrams)
+                probs = data.frame(ngram = bigr_probs$ngram, 
+                                   prob = bigr_probs$prob * 0.4)
+                return(probs)
+        } else {
+                probs = obs_trigs %>% rename( prob = freq) %>%
+                    mutate(prob = prob / bigr_count) %>% 
+                    arrange(desc(prob), ngram)
+                return(probs)
+        }
 
@@
 
-getTrigProbs = function(bigPre, trigrams, bigrams) {
-
+getBigProbs = function(bigPre, bigrams, unigrams) {
+    obs_bigs = getODBigs(bigPre, bigrams)
+    w_i_1 <- str_split(bigPre, "_")[[1]][2]
+    regex = sprintf("%s%s%s", "^", w_i_1, "$")
+    #unig_count = unigrams[grep(regex, unigrams$ngram), ]$freq
+    unig_count = unigrams[ngram %like% regex]$freq
+    if(nrow(obs_bigs) < 1) {
+        probs = getUnigProbs(unigrams) %>% mutate(prob = prob * 0.4)
+        return(probs)
+    } else {
+        probs = obs_bigs %>% rename(prob = freq) %>%
+            mutate(prob = prob/unig_count) %>%
+            arrange(desc(prob), ngram)
+        return(probs)
+    }
 }
 
-S3 = getODTrigs(bigPre, trigs) # get observed trigrams and counts
-S2 = getODBigs(bigPre, bigrs) # get observed bigrams and counts
+getUnigProbs = function(unigrams) {
+    #probs = unigrams %>% rename(prob = freq) %>%
+    #    mutate(prob = prob/sum(prob)) %>%
+    #    arrange(desc(prob), ngram)
+    probs = data.table(unigrams)
+    setnames(probs, "freq", "prob")
+    probs = probs[, prob := prob/sum(prob)][order(prob, decreasing = T)]
+    return(probs)
+}
 

### 2019-11-20 temp.R
benchmarking

@@ -76,3 +76,86 @@ profanity_filter <- function(tv) {
 }
 
 
+
+
+
+
+## DRAFT 1 ---------------------------------------------------------------
+
+library(data.table)
+library(quanteda)
+library(dplyr)
+library(stringr)
+
+getTables = function(dat, ng) {
+        dat.dfm = dfm(dat, ngrams = ng, remove_punct = T, remove_numbers = T)
+        dat.tbl = data.table(ngram = featnames(dat.dfm), freq = colSums(dat.dfm),
+                             key = "ngram")
+        dat.tbl = dat.tbl[order(ngram)]
+        return(dat.tbl)
+}
+
+#set boundaries
+y = list (x1 = 1:350000,
+          x2 = 350001:700000,
+          x3 = 700001:1050000,
+          x4 = 1050001:1400000,
+          x5 = 1400000:1750000,
+          x6 = 1750001:2100000,
+          x7 = 2100001:2360148)
+#read all text
+con <- file("./en_US.twitter.txt", "r") 
+twit = readLines(con, skipNul = T)
+close(con)
+
+train = twit[y[[1]]]
+#transform to corpus
+TW <- corpus(train)
+rm(con, train, twit, x)
+summary(TW, 5)
+
+#merge dfms
+rbindlist(list(a, a2))[, sum(c), b]
+
+TW = corpus(twit[y[[1]]])
+
+unigs1 = getTables(TW, 1)
+bigrs1 = getTables(TW, 2)
+trigs1 = getTables(TW, 3)
+
+
+TW = corpus(twit[y[[2]]])
+
+unigs2 = getTables(TW, 1)
+bigrs2 = getTables(TW, 2)
+trigs2 = getTables(TW, 3)
+
+TW = corpus(twit[y[[3]]])
+
+unigs = rbindlist(list(unigs, unigs1))[, sum(freq), ngram]
+
+
+
+
+
+getFreqs = function(dat, ng) {
+        dat.dfm = dfm(dat, ngrams = ng, remove_punct = T, remove_numbers = T,
+                      remove = stopwords("english"))
+        dat.freq = docfreq(dat.dfm)
+        dat.freq = dat.freq[sort(names(dat.freq))] 
+        return(dat.freq)
+}
+
+getTables2 = function(dat, ng) {
+        ngrams = getFreqs(dat = dat, ng = ng)
[…]

### 2019-11-24 al_stup.Rmd
benchmarking
sqldf package transformation of major functions

@@
 summary(TW, 5)
+
+y = list (x1 = 1:350000,
+          x2 = 350001:700000,
+          x3 = 700001:1050000,
+          x4 = 1050001:1400000,
+          x5 = 1400000:1750000,
+          x6 = 1750001:2100000,
+          x7 = 2100001:2360148)
 ```
@@
 
+getTablesRM = function(dat, ng) {
+    dat.dfm = dfm(dat, ngrams = ng, remove_punct = T, remove_numbers = T,
+                  remove = stopwords("english"))
+    dat.tbl = data.table(ngram = featnames(dat.dfm), freq = colSums(dat.dfm),
+                         key = "ngram")
+    dat.tbl = dat.tbl[order(ngram)]
+    return(dat.tbl)
+}
+
 ```
@@
 trigs = getTables(TW, 3)
+quads = getTables(TW, 4)
 rm(TW)
-
+gc()
 ```
@@
 
+getSQLFinalProbs = function(bigPre, unigs) {
+    querry = paste("SELECT * FROM trigs WHERE ngram like '", 
+                   bigPre, "_%' ORDER BY freq DESC", sep = "")
+    my.dat = as.data.table(sqldf(querry))
+    if(nrow(my.dat) < 1) {
+        my.dat = getSQLBigProbs(bigPre, bigrs)
+        if(nrow(my.dat) < 1) {
+            my.dat = getUnigProbs(unigs)[, freq := freq * 0.4]
+            return(my.dat)
+        } else {
+            my.dat = my.dat[, freq := freq * 0.16]
+            return(my.dat)}
+    } else {
+        total_count = my.dat[, sum(freq)]
+        my.dat = my.dat[ , freq := freq/total_count]
+        return(my.dat)
+    }
+
+}
+
+getSQLTrigProbs = function(bigPre, trigrams) {
+    querry = paste("SELECT * FROM trigs WHERE ngram like '", 
+                   bigPre, "_%' ORDER BY freq DESC", sep = "")
+    my.dat = as.data.table(sqldf(querry))
+    total_count = my.dat[, sum(freq)]
+    my.dat = my.dat[ , freq := freq / total_count][1:50]
+    return(my.dat)
+}
+
+getSQLBigProbs = function(bigPre, bigrams) {
+    w_i_1 <- str_split(bigPre, "_")[[1]][2]
+    querry = paste("SELECT * FROM bigrs WHERE ngram like '",
+                   w_i_1, "_%' ORDER BY freq DESC", sep = "")
+    my.dat = as.data.table(sqldf(querry))
+    total_count = my.dat[, sum(freq)]
+    my.dat = my.dat[, freq := freq/total_count][1:50]
+    return(my.dat)
+}
+
 getUnigProbs = function(unigrams) {
@@
     probs = data.table(unigrams)
-    setnames(probs, "freq", "prob")
-    probs = probs[, prob := prob/sum(prob)][order(prob, decreasing = T)]
+    total_count = probs[, sum(freq)]
+    probs = probs[order(freq, decreasing = T)][1:50][, freq := freq/total_count]
     return(probs)
@@
 
+x = sqldf("SELECT * FROM trigs WHERE ngram like 'settle_the_%' ORDER BY freq DESC")
+
+my.dat = sqldf("SELECT * FROM trigs WHERE ngram like 'how_is_%' ORDER BY freq DESC")
+
+mbm <- microbenchmark(
+    "a" = {x <- getTrigProbs(bigPre, trigs, bigrs, unigs)},
+    "b" = {y <- getSQLFinalProbs(bigPre, trigs)},
+    times = 50
+)
+
 ```
+
+test
+
+```{r}
+#performance test with unobserved regex
+bigPre = "kalem_kagit"
+mbm <- microbenchmark(
+    "a" = {x <- getTrigProbs(bigPre, trigs, bigrs, unigs)},
+    "b" = {y <- getSQLFinalProbs(bigPre, trigs)},
+    times = 50
+)
+mbm
+```
+
+```{r}
+#performance test with observed regex
+bigPre = "how_is"
+mbm <- microbenchmark(
+    "a" = {x <- getTrigProbs(bigPre, trigs, bigrs, unigs)},
+    "b" = {y <- getSQLFinalProbs(bigPre, trigs)},
+    times = 50
+)
+mbm
+```

### 2019-11-24 temp.R

@@ -132,25 +132,42 @@ trigs2 = getTables(TW, 3)
 
 TW = corpus(twit[y[[3]]])
 
-unigs = rbindlist(list(unigs, unigs1))[, sum(freq), ngram]
+unigs3 = getTables(TW, 1)
+bigrs3 = getTables(TW, 2)
+trigs3 = getTables(TW, 3)
 
+TW = corpus(twit[y[[4]]])
 
+unigs4 = getTables(TW, 1)
+bigrs4 = getTables(TW, 2)
+trigs4 = getTables(TW, 3)
 
+TW = corpus(twit[y[[5]]])
 
+unigs5 = getTables(TW, 1)
+bigrs5 = getTables(TW, 2)
+trigs5 = getTables(TW, 3)
 
-getFreqs = function(dat, ng) {
-        dat.dfm = dfm(dat, ngrams = ng, remove_punct = T, remove_numbers = T,
-                      remove = stopwords("english"))
-        dat.freq = docfreq(dat.dfm)
-        dat.freq = dat.freq[sort(names(dat.freq))] 
-        return(dat.freq)
-}
+TW = corpus(twit[y[[6]]])
 
-getTables2 = function(dat, ng) {
-        ngrams = getFreqs(dat = dat, ng = ng)
-        ngrams_dt = data.table(ngram = names(ngrams), freq = ngrams)[order(ngram, decreasing = T)]
-        return(ngrams_dt)
-}
+unigs6 = getTables(TW, 1)
+bigrs6 = getTables(TW, 2)
+trigs6 = getTables(TW, 3)
+
+TW = corpus(twit[y[[7]]])
+
+unigs7 = getTables(TW, 1)
+bigrs7 = getTables(TW, 2)
+trigs7 = getTables(TW, 3)
+
+unigs = rbindlist(list(unigs1, unigs2, unigs3, 
+                       unigs4, unigs5, unigs6, unigs7))[, sum(freq), ngram]
+
+bigrs = rbindlist(list(bigrs1, bigrs2, bigrs3, 
+                       bigrs4, bigrs5, bigrs6, bigrs7))[, sum(freq), ngram]
+
+trigs = rbindlist(list(trigs1, trigs2, trigs3,
+                       trigs4, trigs5, trigs6, trigs7))[, sum(freq), ngram]
 
 
 
[…]

### 2019-11-25 NLP/ui.R
shiny app start

@@ -1,0 +1,43 @@
+#
+# This is the user-interface definition of a Shiny web application. You can
+# run the application by clicking 'Run App' above.
+#
+# Find out more about building applications with Shiny here:
+#
+#    http://shiny.rstudio.com/
+#
+
+library(shiny)
+
+# Define UI for application that draws a histogram
+shinyUI(fluidPage(
+
+    # Application title
+    titlePanel("Word Prediction Using Ngrams"),
+
+    # Sidebar with a slider input for number of bins
+    sidebarLayout(
+        sidebarPanel(
+            textInput("InputText", "Type below:"),
+            submitButton("Submit")
+        ),
+
+        # Show a plot of the generated distribution
+        mainPanel(
+            tabsetPanel(type = "tabs", 
+                        tabPanel("Prediction", 
+                                 textOutput("Associated Words"),
+                                 plotOutput("plot1"),
+                                 h3("Predicted Next Word(s):"),
+                                 textOutput("pred1"),
+                                 textOutput("pred2"),
+                                 textOutput("pred3"),
+                                 textOutput("pred4"),
+                                 textOutput("pred5")
+                        ),
+                        tabPanel("Documentation",
+                                 htmlOutput("documentation"))
+            )
+        )
+    )
+))

### 2019-11-25 al_stup.Rmd
stupid backoff function correction minor additions to server.R

@@
 
+#to get the second word of each row
+getAdjNgram = function(my.ngram) {
+    samp = strsplit(my.ngram[1,ngram], split = "_")
+    if(length(samp[[1]]) == 3) {
+        my.ngram$ngram = str_split_fixed(my.ngram$ngram, "_", 3)[, 3]
+        return(my.ngram)
+    } else if(length(samp[[1]]) == 2) {
+        my.ngram$ngram = str_split_fixed(my.ngram$ngram, "_", 2)[, 2]
+        return(my.ngram)
+    } else return(my.ngram)
+}
+
+#first_words <- str_split_fixed(obsBoBigrams$ngram, "_", 2)[, 1]
+
 getSQLFinalProbs = function(bigPre, unigs) {
+    ID = integer()
     querry = paste("SELECT * FROM trigs WHERE ngram like '", 
@@
         if(nrow(my.dat) < 1) {
-            my.dat = getUnigProbs(unigs)[, freq := freq * 0.4]
+            my.dat = getUnigProbs(unigs)[, freq := freq * 0.16]
             return(my.dat)
         } else {
-            my.dat = my.dat[, freq := freq * 0.16]
+            my.dat = my.dat[, freq := freq * 0.4]
             return(my.dat)}
     } else {
-        total_count = my.dat[, sum(freq)]
-        my.dat = my.dat[ , freq := freq/total_count]
         return(my.dat)
     }
-
 }
@@
     my.dat = as.data.table(sqldf(querry))
-    total_count = my.dat[, sum(freq)]
-    my.dat = my.dat[ , freq := freq / total_count][1:50]
+    w_i_1 <- str_split(bigPre, "_")[[1]][2]
+    querry = paste("SELECT * FROM bigrs WHERE ngram like '",
+                   w_i_1, "_%' ORDER BY freq DESC", sep = "")
+    my.dat.bigs = as.data.table(sqldf(querry))
+    total_count = my.dat.bigs[, sum(freq)]
+    my.dat = my.dat[ , freq := freq / total_count]
     return(my.dat)
@@
     my.dat = as.data.table(sqldf(querry))
-    total_count = my.dat[, sum(freq)]
-    my.dat = my.dat[, freq := freq/total_count][1:50]
+    querry = paste("SELECT * FROM unigsFull WHERE ngram like '",
+                   w_i_1, "' ORDER BY freq DESC", sep = "")
+    my.dat.unig = as.data.table(sqldf(querry))
+    total_count = my.dat.unig[, sum(freq)]
+    my.dat = my.dat[, freq := freq/total_count]
     return(my.dat)

### 2019-11-25 NLP/server.R

@@ -22,5 +22,22 @@ shinyServer(function(input, output) {
         hist(x, breaks = bins, col = 'darkgray', border = 'white')
 
     })
+    
+    output$documentation <- renderUI({
+        str1 <- "Data Source: Credit Card Balance Data (accessed using R Package ISLR)"
+        str2 <- ""
+        str3 <- "Input 1: Select balance value"
+        str4 <- "Input 2: Select rating value"
+        str5 <- "Input 3: Select student status"
+        str6 <- ""
+        str7 <- "Output: Predicted Income level of the person based on unput values"
+        str8 <- ""
+        str9 <- "ui.R code: https://github.com/yigitozanberk/Developing_Data_Products/blob/master/DDP_Final/ui.R"
+        str10 <- ""
+        str11 <- "server.R code: https://github.com/yigitozanberk/Developing_Data_Products/blob/master/DDP_Final/server.R"
+        str12 <- ""
+        HTML(paste(str1, str2, str3, str4, str5, str6, str7, str8, str9, str10, str11, 
+                   str12, sep = '<br/>'))
+    })
 
 })

### 2019-11-25 temp.R

@@ -76,9 +76,20 @@ profanity_filter <- function(tv) {
 }
 
 
+con <- file("./en_US.twitter.txt", "r") 
+twit = readLines(con, skipNul = T)
+close(con)
 
+#takin 15% of data for memory reasons
+set.seed(1)
+x = sample(2360148, 3500, replace = F)
 
+train = twit[x]
 
+TW <- corpus(train)
+
+bigrs = getTables(TW, 2)
+trigs = getTables(TW, 3)
 
 ## DRAFT 1 ---------------------------------------------------------------
 
@@ -86,6 +97,7 @@ library(data.table)
 library(quanteda)
 library(dplyr)
 library(stringr)
+library(sqldf)
 
 getTables = function(dat, ng) {
         dat.dfm = dfm(dat, ngrams = ng, remove_punct = T, remove_numbers = T)

### 2019-11-26 NLP/getSQLTrigProbs.R
shiny app

@@ -1,0 +1,20 @@
+## Returns a two column data.table of predicted trigrams that are the
+## acquired by using the algorithm. If the entered data.table is empty
+## and the input expression is not observed, returns an empty data.table
+##
+## bigPre - cleaned user input.
+## unigs - 2 column data.table. The first column : ngram, contains all 
+##            the unigrams observed in the training corpus
+
+getSQLTrigProbs = function(bigPre, trigrams) {
+        querry = paste("SELECT * FROM trigs WHERE ngram like '", 
+                       bigPre, "_%' ORDER BY freq DESC", sep = "")
+        my.dat = as.data.table(sqldf(querry))
+        w_i_1 <- str_split(bigPre, "_")[[1]][2]
+        querry = paste("SELECT * FROM bigrs WHERE ngram like '",
+                       w_i_1, "_%' ORDER BY freq DESC", sep = "")
+        my.dat.bigs = as.data.table(sqldf(querry))
+        total_count = my.dat.bigs[, sum(freq)]
+        my.dat = my.dat[ , freq := freq / total_count]
+        return(my.dat)
+}

### 2019-11-26 NLP/server.R

@@ -8,30 +8,55 @@
 #
 
 library(shiny)
+library(quanteda)
+library(dplyr)
+library(stringr)
+library(wordcloud)
+library(data.table)
 
 # Define server logic required to draw a histogram
 shinyServer(function(input, output) {
 
-    output$distPlot <- renderPlot({
+    ## load the ngram databases extracted from twitter corpus
+    load("unigsFull.rda")
+    load("unigs.rda")
+    load("bigrs.rda")
+    load("trigs.rda")
+    source("getAdjNgram.R")
+    source("getSQLFinalProbs.R")
+    source("getSQLTrigProbs.R")
+    source("getSQLBigProbs.R")
+    source("getUnigProbs.R")
+    source("wordSplitter.R")
+    
 
-        # generate bins based on input$bins from ui.R
-        x    <- faithful[, 2]
-        bins <- seq(min(x), max(x), length.out = input$bins + 1)
+    modelpred <- reactive({
+        in.words <- input$InputText
+        bigPre <- wordSplitter(in.words)
+        pred <- getSQLFinalProbs(bigPre, unigs)
+        pred.words <- getAdjNgram(pred)
+        return(pred.words)
+    })
+    
+    output$pred1 <- renderText({
+        modelpred()[1]
+    })
+    
+    
+    output$wordPlot <- renderPlot({
 
-        # draw the histogram with the specified number of bins
-        hist(x, breaks = bins, col = 'darkgray', border = 'white')
 
     })
     
     output$documentation <- renderUI({
-        str1 <- "Data Source: Credit Card Balance Data (accessed using R Package ISLR)"
+        str1 <- "Data Source: Coursera Capstone Project Twitter Data"
         str2 <- ""
-        str3 <- "Input 1: Select balance value"
-        str4 <- "Input 2: Select rating value"
-        str5 <- "Input 3: Select student status"
+        str3 <- "Input: Type any word(s)"
+        str4 <- ""
+        str5 <- ""
         str6 <- ""
-        str7 <- "Output: Predicted Income level of the person based on unput values"
-        str8 <- ""
+        str7 <- "Output: Top 5 predicted words"
+        str8 <- "Output: Top associated words as wordcloud image"
         str9 <- "ui.R code: https://github.com/yigitozanberk/Developing_Data_Products/blob/master/DDP_Final/ui.R"
         str10 <- ""
         str11 <- "server.R code: https://github.com/yigitozanberk/Developing_Data_Products/blob/master/DDP_Final/server.R"

### 2019-11-26 NLP/ui.R

@@ -27,13 +27,14 @@ shinyUI(fluidPage(
             tabsetPanel(type = "tabs", 
                         tabPanel("Prediction", 
                                  textOutput("Associated Words"),
-                                 plotOutput("plot1"),
                                  h3("Predicted Next Word(s):"),
+                                 h3("(Please wait while the program)"),
                                  textOutput("pred1"),
                                  textOutput("pred2"),
                                  textOutput("pred3"),
                                  textOutput("pred4"),
-                                 textOutput("pred5")
+                                 textOutput("pred5"),
+                                 plotOutput("plot1"),
                         ),
                         tabPanel("Documentation",
                                  htmlOutput("documentation"))

### 2019-11-26 NLP/wordSplitter.R

@@ -1,0 +1,18 @@
+## cleans the input text into required format and outputs text
+## in the form w1_w2_w3, w1_w1, or w1. 
+## Maximum format is with 3 words.
+## 
+## inputWords - user input text
+wordSplitter = function(inputWords) {
+        s.words = unlist(strsplit(inputWords, "[ !?.]+"))
+        count = length(s.words)
+        bigPre = character()
+        if(count >= 2) {
+                bigPre = paste(s.words[count-1],
+                               s.words[count], sep = "_")
+                return(bigPre)
+        } else {
+                bigPre = s.words
+                return(bigPre)
+        }
+}

### 2019-11-26 backup.R

@@ -1,0 +1,120 @@
+## Returns a two column data.table of predicted ngrams that are the
+## acquired by using the algorithm. If the entered data.table is empty
+## then an empty data.table is returned.
+##
+## my.ngram - 2 column data.table. The first column: ngram,
+##            contains all the predicted trigrams, bigrams or unigrams
+##            together with their associated probability scores
+getAdjNgram = function(my.ngram) {
+        samp = strsplit(my.ngram[1,ngram], split = "_")
+        if(length(samp[[1]]) == 3) {
+                my.ngram$ngram = str_split_fixed(my.ngram$ngram, "_", 3)[, 3]
+                return(my.ngram)
+        } else if(length(samp[[1]]) == 2) {
+                my.ngram$ngram = str_split_fixed(my.ngram$ngram, "_", 2)[, 2]
+                return(my.ngram)
+        } else return(my.ngram)
+}
+
+## Returns a two column data.table of predicted ngrams that are the
+## acquired by using the algorithm. If the entered data.table is empty
+## and the input expression is not observed, returns an empty data.table
+##
+## bigPre - cleaned user input.
+## unigs - 2 column data.table. The first column : ngram, contains all 
+##            the unigrams observed in the training corpus
+
+getSQLFinalProbs = function(bigPre, unigs) {
+        ID = integer()
+        querry = paste("SELECT * FROM trigs WHERE ngram like '", 
+                       bigPre, "_%' ORDER BY freq DESC", sep = "")
+        my.dat = as.data.table(sqldf(querry))
+        if(nrow(my.dat) < 1) {
+                my.dat = getSQLBigProbs(bigPre, bigrs)
+                if(nrow(my.dat) < 1) {
+                        my.dat = getUnigProbs(unigs)[, freq := freq * 0.16]
+                        return(my.dat)
+                } else {
+                        my.dat = my.dat[, freq := freq * 0.4]
+                        return(my.dat)}
+        } else {
+                return(my.dat)
+        }
+}
+
+## Returns a two column data.table of predicted trigrams that are the
+## acquired by using the algorithm. If the entered data.table is empty
+## and the input expression is not observed, returns an empty data.table
+##
+## bigPre - cleaned user input.
+## trigrams - 2 column data.table. The first column : ngram, contains all 
+##            the unigrams observed in the training corpus
+
[…]

### 2019-11-26 getAdjNgram.R

@@ -1,0 +1,17 @@
+## Returns a two column data.table of predicted ngrams that are the
+## acquired by using the algorithm. If the entered data.table is empty
+## then an empty data.table is returned.
+##
+## my.ngram - 2 column data.table. The first column: ngram,
+##            contains all the predicted trigrams, bigrams or unigrams
+##            together with their associated probability scores
+getAdjNgram = function(my.ngram) {
+        samp = strsplit(my.ngram[1,ngram], split = "_")
+        if(length(samp[[1]]) == 3) {
+                my.ngram$ngram = str_split_fixed(my.ngram$ngram, "_", 3)[, 3]
+                return(my.ngram)
+        } else if(length(samp[[1]]) == 2) {
+                my.ngram$ngram = str_split_fixed(my.ngram$ngram, "_", 2)[, 2]
+                return(my.ngram)
+        } else return(my.ngram)
+}

### 2019-11-26 getSQLBigProbs.R

@@ -1,0 +1,19 @@
+## Returns a two column data.table of predicted bigrams that are the
+## acquired by using the algorithm. If the entered data.table is empty
+## and the input expression is not observed, returns an empty data.table
+##
+## bigPre - cleaned user input.
+## unigs - 2 column data.table. The first column : ngram, contains all 
+##            the unigrams observed in the training corpus
+getSQLBigProbs = function(bigPre, bigrams) {
+        w_i_1 <- str_split(bigPre, "_")[[1]][2]
+        querry = paste("SELECT * FROM bigrs WHERE ngram like '",
+                       w_i_1, "_%' ORDER BY freq DESC", sep = "")
+        my.dat = as.data.table(sqldf(querry))
+        querry = paste("SELECT * FROM unigsFull WHERE ngram like '",
+                       w_i_1, "' ORDER BY freq DESC", sep = "")
+        my.dat.unig = as.data.table(sqldf(querry))
+        total_count = my.dat.unig[, sum(freq)]
+        my.dat = my.dat[, freq := freq/total_count]
+        return(my.dat)
+}

### 2019-11-26 getSQLFinalProbs.R

@@ -1,0 +1,25 @@
+## Returns a two column data.table of predicted ngrams that are the
+## acquired by using the algorithm. If the entered data.table is empty
+## and the input expression is not observed, returns an empty data.table
+##
+## bigPre - cleaned user input.
+## unigs - 2 column data.table. The first column : ngram, contains all 
+##            the unigrams observed in the training corpus
+
+getSQLFinalProbs = function(bigPre, unigs) {
+        ID = integer()
+        querry = paste("SELECT * FROM trigs WHERE ngram like '", 
+                       bigPre, "_%' ORDER BY freq DESC", sep = "")
+        my.dat = as.data.table(sqldf(querry))
+        if(nrow(my.dat) < 1) {
+                my.dat = getSQLBigProbs(bigPre, bigrs)
+                if(nrow(my.dat) < 1) {
+                        my.dat = getUnigProbs(unigs)[, freq := freq * 0.16]
+                        return(my.dat)
+                } else {
+                        my.dat = my.dat[, freq := freq * 0.4]
+                        return(my.dat)}
+        } else {
+                return(my.dat)
+        }
+}

### 2019-11-26 getUnigProbs.R

@@ -1,0 +1,15 @@
+## Returns a two column data.table of predicted unigrams that are the
+## acquired by using the algorithm. If the entered data.table is empty
+## and the input expression is not observed, returns an empty data.table
+##
+## unigs - 2 column data.table. The first column : ngram, contains all 
+##            the unigrams observed in the training corpus
+getUnigProbs = function(unigrams) {
+        #probs = unigrams %>% rename(prob = freq) %>%
+        #    mutate(prob = prob/sum(prob)) %>%
+        #    arrange(desc(prob), ngram)
+        probs = data.table(unigrams)
+        total_count = probs[, sum(freq)]
+        probs = probs[order(freq, decreasing = T)][1:50][, freq := freq/total_count]
+        return(probs)
+}

### 2019-11-26 temp.R

@@ -90,6 +90,7 @@ TW <- corpus(train)
 
 bigrs = getTables(TW, 2)
 trigs = getTables(TW, 3)
+unigs = getTables(TW, 1)
 
 ## DRAFT 1 ---------------------------------------------------------------
 

### 2019-11-28 NLP/server.R
shiny app done

@@ -13,22 +13,23 @@ library(dplyr)
 library(stringr)
 library(wordcloud)
 library(data.table)
+library(sqldf)
 
 # Define server logic required to draw a histogram
 shinyServer(function(input, output) {
 
     ## load the ngram databases extracted from twitter corpus
-    load("unigsFull.rda")
-    load("unigs.rda")
-    load("bigrs.rda")
-    load("trigs.rda")
+    #load("unigsFull.rda")
+    #load("unigs.rda")
+    #load("bigrs.rda")
+    #load("trigs.rda")
     source("getAdjNgram.R")
     source("getSQLFinalProbs.R")
     source("getSQLTrigProbs.R")
     source("getSQLBigProbs.R")
     source("getUnigProbs.R")
     source("wordSplitter.R")
-    
+    mypred <- vector(mode = "character", length = 0 )
 
     modelpred <- reactive({
         in.words <- input$InputText
@@ -39,15 +40,31 @@ shinyServer(function(input, output) {
     })
     
     output$pred1 <- renderText({
-        modelpred()[1]
+        modelpred()[1, ngram]
+    })
+    output$pred2 <- renderText({
+        modelpred()[2, ngram]
+    })
+    output$pred3 <- renderText({
+        modelpred()[3, ngram]
+    })
+    output$pred4 <- renderText({
+        modelpred()[4, ngram]
+    })
+    output$pred5 <- renderText({
+        modelpred()[5, ngram]
     })
     
+    # Make the wordcloud drawing predictable during a session
+    wordcloud_rep <- repeatable(wordcloud)
     
-    output$wordPlot <- renderPlot({
-
-
+    output$plot <- renderPlot({
+        mytab <- modelpred()
+        wordcloud_rep(mytab[, ngram], mytab[, freq], scale=c(4,0.5),
+                      min.freq = 0, max.words= 50,
+                      color = RColorBrewer::brewer.pal(8, "Dark2"))
     })
-    
+
     output$documentation <- renderUI({
         str1 <- "Data Source: Coursera Capstone Project Twitter Data"
         str2 <- ""

### 2019-11-28 NLP/ui.R

@@ -28,13 +28,13 @@ shinyUI(fluidPage(
                         tabPanel("Prediction", 
                                  textOutput("Associated Words"),
                                  h3("Predicted Next Word(s):"),
-                                 h3("(Please wait while the program)"),
+                                 h6("(Please wait for the program to initialize in the first run)"),
                                  textOutput("pred1"),
                                  textOutput("pred2"),
                                  textOutput("pred3"),
                                  textOutput("pred4"),
                                  textOutput("pred5"),
-                                 plotOutput("plot1"),
+                                 plotOutput("plot"),
                         ),
                         tabPanel("Documentation",
                                  htmlOutput("documentation"))

### 2019-11-28 backup.R

@@ -118,3 +118,97 @@ wordSplitter = function(inputWords) {
         }
 }
 
+
+
+
+
+
+
+
+### SERVER CODE
+
+#
+# This is the server logic of a Shiny web application. You can run the
+# application by clicking 'Run App' above.
+#
+# Find out more about building applications with Shiny here:
+#
+#    http://shiny.rstudio.com/
+#
+
+library(shiny)
+library(quanteda)
+library(dplyr)
+library(stringr)
+library(wordcloud)
+library(data.table)
+library(sqldf)
+
+# Define server logic required to draw a histogram
+shinyServer(function(input, output) {
+        
+        ## load the ngram databases extracted from twitter corpus
+        #load("unigsFull.rda")
+        #load("unigs.rda")
+        #load("bigrs.rda")
+        #load("trigs.rda")
+        source("getAdjNgram.R")
+        source("getSQLFinalProbs.R")
+        source("getSQLTrigProbs.R")
+        source("getSQLBigProbs.R")
+        source("getUnigProbs.R")
+        source("wordSplitter.R")
+        mypred <- vector(mode = "character", length = 0 )
+        
+        modelpred <- reactive({
+                in.words <- input$InputText
+                bigPre <- wordSplitter(in.words)
+                pred <- getSQLFinalProbs(bigPre, unigs)
+                pred.words <- getAdjNgram(pred)
+                return(pred.words)
+        })
+        
+        output$pred1 <- renderText({
+                modelpred()[1, ngram]
+        })
+        output$pred2 <- renderText({
+                modelpred()[2, ngram]
+        })
+        output$pred3 <- renderText({
+                modelpred()[3, ngram]
+        })
+        output$pred4 <- renderText({
+                modelpred()[4, ngram]
+        })
+        output$pred5 <- renderText({
+                modelpred()[5, ngram]
+        })
+        
+        # Make the wordcloud drawing predictable during a session
+        wordcloud_rep <- repeatable(wordcloud)
+        
+        output$plot <- renderPlot({
+                mytab <- modelpred()
+                wordcloud_rep(mytab[, ngram], mytab[, freq], scale=c(4,0.5),
+                              min.freq = 0, max.words= 50,
+                              color = RColorBrewer::brewer.pal(8, "Dark2"))
+        })
+        
+        output$documentation <- renderUI({
+                str1 <- "Data Source: Coursera Capstone Project Twitter Data"
+                str2 <- ""
[…]

### 2019-11-28 NLP/server.R
progress bar added

@@ -19,10 +19,10 @@ library(sqldf)
 shinyServer(function(input, output) {
 
     ## load the ngram databases extracted from twitter corpus
-    #load("unigsFull.rda")
-    #load("unigs.rda")
-    #load("bigrs.rda")
-    #load("trigs.rda")
+    load("unigsFull.rda")
+    load("unigs.rda")
+    load("bigrs.rda")
+    load("trigs.rda")
     source("getAdjNgram.R")
     source("getSQLFinalProbs.R")
     source("getSQLTrigProbs.R")
@@ -32,11 +32,20 @@ shinyServer(function(input, output) {
     mypred <- vector(mode = "character", length = 0 )
 
     modelpred <- reactive({
-        in.words <- input$InputText
-        bigPre <- wordSplitter(in.words)
-        pred <- getSQLFinalProbs(bigPre, unigs)
-        pred.words <- getAdjNgram(pred)
-        return(pred.words)
+            withProgress({
+                setProgress(message = "Processing corpus...")
+                in.words <- input$InputText
+                bigPre <- wordSplitter(in.words)
+                pred <- getSQLFinalProbs(bigPre, unigs)
+                pred.words <- getAdjNgram(pred)
+                return(pred.words)
+            })
+        
+        #in.words <- input$InputText
+        #bigPre <- wordSplitter(in.words)
+        #pred <- getSQLFinalProbs(bigPre, unigs)
+        #pred.words <- getAdjNgram(pred)
+        #return(pred.words)
     })
     
     output$pred1 <- renderText({

### 2019-11-28 NLP/server.R
added correct hardcode addresses

@@ -83,9 +83,9 @@ shinyServer(function(input, output) {
         str6 <- ""
         str7 <- "Output: Top 5 predicted words"
         str8 <- "Output: Top associated words as wordcloud image"
-        str9 <- "ui.R code: https://github.com/yigitozanberk/Developing_Data_Products/blob/master/DDP_Final/ui.R"
+        str9 <- "ui.R code: https://github.com/yigitozanberk/capstone/blob/master/NLP/ui.R"
         str10 <- ""
-        str11 <- "server.R code: https://github.com/yigitozanberk/Developing_Data_Products/blob/master/DDP_Final/server.R"
+        str11 <- "server.R code: https://github.com/yigitozanberk/capstone/blob/master/NLP/server.R"
         str12 <- ""
         HTML(paste(str1, str2, str3, str4, str5, str6, str7, str8, str9, str10, str11, 
                    str12, sep = '<br/>'))

### 2019-11-30 al_stup.Rmd
shiny app
server code with rds

@@
 library(stringr)
+library(wordcloud)
+library(data.table)
 library(readtext)
@@
 library(ggplot2)
-library(wordcloud)
-library(data.table)
+
 ```
@@
 set.seed(1)
-x = sample(2360148, 350000, replace = F)
+x = sample(2360148, 590037, replace = F)
 

### 2019-11-30 NLP/server.R

@@ -8,8 +8,6 @@
 #
 
 library(shiny)
-library(quanteda)
-library(dplyr)
 library(stringr)
 library(wordcloud)
 library(data.table)
@@ -19,10 +17,10 @@ library(sqldf)
 shinyServer(function(input, output) {
 
     ## load the ngram databases extracted from twitter corpus
-    load("unigsFull.rda")
     load("unigs.rda")
     load("bigrs.rda")
     load("trigs.rda")
+    load("unigsFull.rda")
     source("getAdjNgram.R")
     source("getSQLFinalProbs.R")
     source("getSQLTrigProbs.R")

### 2019-11-30 backup.R

@@ -212,3 +212,14 @@ shinyServer(function(input, output) {
         })
         
 })
+
+
+library()
+unigsFull = readRDS("unigsFull.rds")
+unigs = readRDS("unigs.rds")
+bigrs = readRDS("bigrs.rds")
+trigs = readRDS("trigs.rds")
+#source_data("https://github.com/yigitozanberk/capstone/blob/master/NLP/trigs.rda?raw=true")
+#source_data("https://github.com/yigitozanberk/capstone/blob/master/NLP/bigrs.rda?raw=true")
+#source_data("https://github.com/yigitozanberk/capstone/blob/master/NLP/unigs.rda?raw=true")
+#source_data("https://github.com/yigitozanberk/capstone/blob/master/NLP/unigsFull.rda?raw=true")

### 2019-11-30 al_stup.Rmd
server

@@
 library(ggplot2)
-
+library(sqldf)
 ```
@@
 set.seed(1)
-x = sample(2360148, 590037, replace = F)
+x = sample(2360148, 300000, replace = F)
 
@@
 unigs = getTables(TW, 1)
+unigsfull = getTablesRM(TW, 1)
 bigrs = getTables(TW, 2)
 trigs = getTables(TW, 3)
-quads = getTables(TW, 4)
 rm(TW)

### 2019-11-30 NLP/server.R

@@ -20,13 +20,21 @@ shinyServer(function(input, output) {
     load("unigs.rda")
     load("bigrs.rda")
     load("trigs.rda")
-    load("unigsFull.rda")
+    load("unigsfull.rda")
+    #unigsFull <- data.table(readRDS("./unigsfull.rds"))
+    #unigs <- data.table(readRDS("./unigs.rds"))
+    #bigrs <- data.table(readRDS("./bigrs.rds"))
+    #trigs <- data.table(readRDS("./trigs.rds"))
     source("getAdjNgram.R")
     source("getSQLFinalProbs.R")
     source("getSQLTrigProbs.R")
     source("getSQLBigProbs.R")
     source("getUnigProbs.R")
     source("wordSplitter.R")
+    trigs <- data.table(trigs)
+    bigrs <- data.table(bigrs)
+    unigs <- data.table(unigs)
+    unigsFull <- data.table(unigsfull)
     mypred <- vector(mode = "character", length = 0 )
 
     modelpred <- reactive({
@@ -34,9 +42,14 @@ shinyServer(function(input, output) {
                 setProgress(message = "Processing corpus...")
                 in.words <- input$InputText
                 bigPre <- wordSplitter(in.words)
-                pred <- getSQLFinalProbs(bigPre, unigs)
-                pred.words <- getAdjNgram(pred)
-                return(pred.words)
+                if(length(bigPre) == 0) {
+                    pred.words = getUnigProbs(unigsFull)[1:50]
+                    return(pred.words)
+                } else {
+                    pred <- getSQLFinalProbs(bigPre, unigs)
+                    pred.words <- getAdjNgram(pred)
+                    return(pred.words)
+                }
             })
         
         #in.words <- input$InputText

### 2019-11-30 al_stup.Rmd
final version

@@
                 bigr_probs = getBigProbs(bigPre, bigrams, unigrams)
-                probs = data.frame(ngram = bigr_probs$ngram, 
+                probs = data.table(ngram = bigr_probs$ngram, 
                                    prob = bigr_probs$prob * 0.4)

### 2019-11-30 NLP/server.R

@@ -12,6 +12,7 @@ library(stringr)
 library(wordcloud)
 library(data.table)
 library(sqldf)
+library(dplyr)
 
 # Define server logic required to draw a histogram
 shinyServer(function(input, output) {
@@ -31,6 +32,10 @@ shinyServer(function(input, output) {
     source("getSQLBigProbs.R")
     source("getUnigProbs.R")
     source("wordSplitter.R")
+    source("getTrigProbs.R")
+    source("getBigProbs.R")
+    source("getODTrigs.R")
+    source("getODBigs.R")
     trigs <- data.table(trigs)
     bigrs <- data.table(bigrs)
     unigs <- data.table(unigs)
@@ -43,10 +48,11 @@ shinyServer(function(input, output) {
                 in.words <- input$InputText
                 bigPre <- wordSplitter(in.words)
                 if(length(bigPre) == 0) {
-                    pred.words = getUnigProbs(unigsFull)[1:50]
+                    pred.words = getUnigProbs(unigsFull)[1:50] %>%
+                        rename(prob = freq)
                     return(pred.words)
                 } else {
-                    pred <- getSQLFinalProbs(bigPre, unigs)
+                    pred <- getTrigProbs(bigPre, trigs, bigrs, unigs)
                     pred.words <- getAdjNgram(pred)
                     return(pred.words)
                 }
@@ -80,7 +86,7 @@ shinyServer(function(input, output) {
     
     output$plot <- renderPlot({
         mytab <- modelpred()
-        wordcloud_rep(mytab[, ngram], mytab[, freq], scale=c(4,0.5),
+        wordcloud_rep(mytab[, ngram], mytab[, prob], scale=c(4,0.5),
                       min.freq = 0, max.words= 50,
                       color = RColorBrewer::brewer.pal(8, "Dark2"))
     })

### 2019-11-30 backup.R

@@ -223,3 +223,13 @@ trigs = readRDS("trigs.rds")
 #source_data("https://github.com/yigitozanberk/capstone/blob/master/NLP/bigrs.rda?raw=true")
 #source_data("https://github.com/yigitozanberk/capstone/blob/master/NLP/unigs.rda?raw=true")
 #source_data("https://github.com/yigitozanberk/capstone/blob/master/NLP/unigsFull.rda?raw=true")
+
+
+
+
+## https://yoberk.shinyapps.io/mypred/
+        
+        
+        
+        
+        

### 2019-11-30 getAdjNgram.R

@@ -6,6 +6,7 @@
 ##            contains all the predicted trigrams, bigrams or unigrams
 ##            together with their associated probability scores
 getAdjNgram = function(my.ngram) {
+        my.ngram = data.table(my.ngram)
         samp = strsplit(my.ngram[1,ngram], split = "_")
         if(length(samp[[1]]) == 3) {
                 my.ngram$ngram = str_split_fixed(my.ngram$ngram, "_", 3)[, 3]

### 2019-12-02 al_stup.Rmd
presentation

@@
         bigr_count = bigrams[ngram %like% regex]$freq
-
-        if(nrow(obs_trigs) < 1) {
+        count = length(strsplit(bigPre, "_")[[1]])
+        if(count < 2 | nrow(obs_trigs) < 1) {
                 bigr_probs = getBigProbs(bigPre, bigrams, unigrams)
@@
     if(nrow(obs_bigs) < 1) {
-        probs = getUnigProbs(unigrams) %>% mutate(prob = prob * 0.4)
+        probs = getUnigProbs(unigrams) %>% mutate(prob = freq * 0.4)
         return(probs)

### 2019-12-02 NLP/server.R

@@ -39,7 +39,7 @@ shinyServer(function(input, output) {
     trigs <- data.table(trigs)
     bigrs <- data.table(bigrs)
     unigs <- data.table(unigs)
-    unigsFull <- data.table(unigsfull)
+    unigsfull <- data.table(unigsfull)
     mypred <- vector(mode = "character", length = 0 )
 
     modelpred <- reactive({
@@ -48,7 +48,7 @@ shinyServer(function(input, output) {
                 in.words <- input$InputText
                 bigPre <- wordSplitter(in.words)
                 if(length(bigPre) == 0) {
-                    pred.words = getUnigProbs(unigsFull)[1:50] %>%
+                    pred.words = getUnigProbs(unigsfull)[1:50] %>%
                         rename(prob = freq)
                     return(pred.words)
                 } else {

### 2019-12-02 backup.R

@@ -228,7 +228,7 @@ trigs = readRDS("trigs.rds")
 
 
 ## https://yoberk.shinyapps.io/mypred/
-        
+## http://rpubs.com/yoberk/nlpshinyapp
         
         
         
